---
title: "Lesson 4: Simultaneous Games II: Continuous Strategies"
subtitle: "ECON 316 - Game Theory - Spring 2019"
author: "Ryan Safner"
date: ""
output: 
  beamer_presentation:
    incremental: true 
    fig_caption: yes
    toc: true 
    slide_level: 3
    includes:
      in_header: ../../../templates/template-lecture-slides.tex
    keep_tex: no 
    latex_engine: xelatex #required for Fira Sans font to render properly 
classoption: aspectratio=169
---

```{r setup, include=FALSE}
library(tidyverse)
library(mosaic)
update_geom_defaults("label", list(family = "Yanone Kaffeesatz Regular"))
# For making transparent ggplot2 graphs, from https://gist.github.com/cboettig/5600558 

# Set plotting to bw plot default, but with transparent background elements.  
# Note transparency requires the panel.background, plot.background, and device background all be set!
theme_set(theme_bw(base_size = 9))
theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
             plot.background = element_rect(fill = "transparent", colour = NA))
knitr::opts_chunk$set(echo = F, warning = F, message = F,dev.args=list(bg="transparent"))

```

```{r}
extrafont::loadfonts()
```

### Discrete vs. Continuous Strategies

\columnsbegin

\column{.7\textwidth}

- We considered simultaneous games where strategies are **discrete**: a finite (usually small) number of strategies to choose from
- We now consider **continuous** strategies:
    - Practically infinite number of "strategies" to choose from
    - Example: firms don't just choose 1 out of 2 or 3 possible prices, but one from a large possible range
    - Must switch from game trees and payoff matrices to algebraic equations
    - Still keep concepts of "best response" and Nash equilibria

\column{.3\textwidth}

\begin{figure}
  \includegraphics[height=1.5in]{continuous}
\end{figure}

\columnsend

# Example: The Economics of Oligopoly

### Quick Microeconomics Refresher

\columnsbegin

\column{.5\textwidth}

\begin{tikzpicture}\scriptsize 
	\begin{axis}[
		scale=0.8,
		axis lines=middle, 
		enlarge x limits={rel=0.1, upper},
		enlarge y limits={rel=0.1, upper},
		every axis y label/.style={at={(axis description cs:-0.2,0.5)},rotate=90,anchor=north},
		every axis x label/.style={at={(axis description cs:0.5,-0.1)},anchor=north},
	legend style={at={(0.5,-0.25)},anchor=north},
	xlabel=Output (q),
	ylabel=Price (p),
	xtick={1, 2,...,10},
	ytick={0,5,10,...,50},
	grid=major,
	ymin=0,
	xmin=0,
	ymax=50,
	xmax=10,
	legend cell align=left,
]
	\addplot[domain=0:10,ultra thick,blue!50]{50-2*x};
	\addlegendentry{Demand}
	\only<4->{\addplot[domain=0:10,ultra thick,purple!75]{50-4*x};
	\addlegendentry{$MR(q)$}}
	\only<5->{\addplot[domain=0:10,ultra thick, red!50]{30};
	\addlegendentry{$MC(q)=AC(q)$}}
	\draw<6->[fill=black] (axis cs:5,40)circle(0.075cm)node[above]{$\pi$-max};
	\draw<6->[dashed, thick, black] (axis cs:0,40)node[right=.25cm, above]{$p^*$}--(axis cs:5,40)--(axis cs:5,0)node[above=.25cm, right]{$q^*$};
\only<7->{\addplot[fill=teal, fill opacity = 0.2, draw = none,area legend] coordinates {(0,30) (0,40) (5,40) (5,30)};
       \addlegendentry{Profit $\pi$};}
        \end{axis}
\end{tikzpicture}	

\column{.5\textwidth}

- We often represent demand as a linear function
$$p=a-bq$$
    - $a$: vertical intercept ("choke price")
    - $b$: slope
- \textcolor{purple}{Marginal revenue} curve is twice the slope of demand curve
- Simplify with a constant \textcolor{red}{$MC(q)$}, no fixed costs, thus $MC(q)=AC(q)$
- Firms maximize profit at $q^*$ where $\textcolor{purple}{MR(q)}=\textcolor{red}{MC(q)}$, raise price to highest consumers are willing to pay (\textcolor{blue}{demand})
- \textcolor{teal}{Total profit} $\textcolor{teal}{\pi}=q^*(p^*-AC)$

\columnsend

### Monopoly vs. Perfect Competition I

\columnsbegin

\column{.5\textwidth}

\begin{tikzpicture}\scriptsize 
	\begin{axis}[
		scale=0.8,
		axis lines=middle, 
		enlarge x limits={rel=0.1, upper},
		enlarge y limits={rel=0.1, upper},
		every axis y label/.style={at={(axis description cs:-0.2,0.5)},rotate=90,anchor=north},
		every axis x label/.style={at={(axis description cs:0.5,-0.1)},anchor=north},
	xlabel=Output (q),
	ylabel=Price (p),
	xtick={1, 2,...,10},
	ytick={0,5,10,...,50},
	legend style={at={(0.5,-0.25)},anchor=north},
	grid=major,
	ymin=0,
	xmin=0,
	ymax=50,
	xmax=10,
	legend cell align=left,
]
	\addplot[domain=0:10,ultra thick, red!50]{5*x};
	\addlegendentry{$MC(q)$}
	\addplot[domain=0:10,ultra thick, orange!50]{2.5*x+10/x};
	\addlegendentry{$AC(q)$}
	\addplot[domain=0:10,ultra thick,blue!50]{10};
	\addlegendentry{Demand $=p=MR(q)$}
	\draw[fill=black] (axis cs:2,10)circle(0.075cm)node[above]{$\pi$-max};
	\draw[dashed, thick, black] (axis cs:0,10)node[right=.25cm, above]{$p^*$}--(axis cs:2,10)--(axis cs:2,0)node[above=.25cm, right]{$q^*$};
        \end{axis}
\end{tikzpicture}	

\column{.5\textwidth}

- In **perfect competition**, large number of firms selling perfect substitutes
- Firms are "**price-takers**", demand for firm is perfectly elastic at exogenous market price
    - Demand$=$MR$=$price
    - $q^*$ where $MR(q)=MC(q)$
    - $p^*=MC$
- Long run profits $\pi=0$ due to **free entry** and exit

\columnsend 

### Monopoly vs. Perfect Competition II 

\columnsbegin

\column{.5\textwidth}

\begin{tikzpicture}\scriptsize 
	\begin{axis}[
		scale=0.8,
		axis lines=middle, 
		enlarge x limits={rel=0.1, upper},
		enlarge y limits={rel=0.1, upper},
		every axis y label/.style={at={(axis description cs:-0.2,0.5)},rotate=90,anchor=north},
		every axis x label/.style={at={(axis description cs:0.5,-0.1)},anchor=north},
	xlabel=Output (q),
	ylabel=Price (p),
	xtick={1, 2,...,10},
	ytick={0,5,10,...,50},
	legend style={at={(0.5,-0.25)},anchor=north},
	grid=major,
	ymin=0,
	xmin=0,
	ymax=50,
	xmax=10,
	legend cell align=left,
]
	\addplot[domain=0:10,ultra thick, red!50]{5*x};
	\addlegendentry{$MC(q)$}
	\addplot[domain=0:10,ultra thick, orange!50]{2.5*x+10/x};
	\addlegendentry{$AC(q)$}
	\addplot[domain=0:10,ultra thick,blue!50]{50-2.5*x};
	\addlegendentry{Demand}
	\addplot[domain=0:10,ultra thick,purple!50]{50-5*x};
	\addlegendentry{$MR(q)$}
	\draw[fill=black] (axis cs:5,37.5)circle(0.075cm)node[above]{$\pi$-max};
	\draw[dashed, thick, black] (axis cs:0,37.5)node[right=.25cm, above]{$p^*$}--(axis cs:5,37.5)--(axis cs:5,0)node[above=.25cm, right]{$q^*$};
\addplot[fill=teal, fill opacity = 0.2, draw = none,area legend] coordinates {(0,15) (0,37.5) (5,37.5) (5,15)};
\addlegendentry{Profit $\pi$};
        \end{axis}
\end{tikzpicture}	

\column{.5\textwidth}

- In **monopoly**, a single firm, faces entire market demand curve
- A **price-searcher**, must pick both $q^*$ and $p^*$
    - $q^*$ where $MR(q)=MC(q)$
    - $p^*$: market demand at $q^*$
- Can earn long run profits $\pi$ due to **barriers to entry**

\columnsend

### Oligopoly

\begin{figure}
		\includegraphics[height=1.75in]{wirelesscarriers}
	  \includegraphics[height=1.75in]{big6studios}	
\end{figure}

### Oligopoly: Features/Assumptions

- \alert{Oligopoly}: an industry with a few large sellers with market power
- **Duopoly**: a market with only two sellers; *Triopoly*: 3 sellers, etc. 
- Firms make *strategic choices*, interdependent on one another 
- May be selling the same good, meeting one market demand 
- Often barriers to entry

### Oligopoly II

\columnsbegin

\column{.7\textwidth}

- Unlike PC or monopoly, no single "theory of oligopoly"
- Depends heavily on assumptions made about interactions and choice variables
- One certainty: oligopoly is a strategic interaction between few firms (i.e. ideal for game theory)

\column{.3\textwidth}

\begin{figure}
  \includegraphics[height=1.5in]{cokevspepsi}
\end{figure}

\columnsend

### Equilibrium in Oligopoly

\columnsbegin

\column{.7\textwidth}

- What does "**equilibrium**" mean in an oligopoly?
- In PC/monopoly, a vector of $(q^*,p^*)$ for whole market such that no firms/consumers have incentives to change price
- Oligopoly: we can use **Nash Equilibrium**: 
    - no firm wants to change what they are doing given all others' strategies
    - each firm is playing a **best response** against other firms' strategies

\column{.3\textwidth}

\begin{figure}
  \includegraphics[height=1.125in]{equilibriumbalance}
\end{figure}

\columnsend

## As Prisoners' Dilemma

### Oligopoly as Simple Prisoners' Dilemma

- Suppose we have a duopoly between \textcolor{red}{Apple} and \textcolor{blue}{Google}
- Each is planning to launch a new tablet, and choose to sell it at a High Price or a Low Price	

\begin{figure}
\centering 
\begin{tabular}{lrcc|cc|}
& & \multicolumn{4}{c} {\textbf{\textcolor{blue}{Google}}} \\
& & \multicolumn{2}{c}{High Price} & \multicolumn{2}{c}{Low Price} \\ \cline{3-6}
\parbox[t]{2mm}{\multirow{4}{*}{\rotatebox[origin=c]{90}{\textbf{\textcolor{red}{Apple}}}}}  & 
\multicolumn{1}{c|}{\multirow{2}{*}{High Price}} & \$500,000 & & \$100,000 & \\ 
&  &\multicolumn{1}{|c}{} & \$500,000 & & \$750,000\\ \cline{3-6}
&
\multicolumn{1}{c|}{\multirow{2}{*}{Low Price}} & \$750,000 &  & \$250,000 & \\ 
& & \multicolumn{1}{|c}{} & \$100,000 &  & \$250,000 \\ \cline{3-6}
\end{tabular}
\caption*{Payoffs represent profits}  
\end{figure} 

### Oligopoly as Simple Prisoners' Dilemma II

- Low Price is a dominant strategy for \textcolor{red}{Apple}

\begin{figure}
\centering 
\begin{tabular}{lrcc|cc|}
& & \multicolumn{4}{c} {\textbf{\textcolor{blue}{Google}}} \\
& & \multicolumn{2}{c}{High Price} & \multicolumn{2}{c}{Low Price} \\ \cline{3-6}
\parbox[t]{2mm}{\multirow{4}{*}{\rotatebox[origin=c]{90}{\textbf{\textcolor{red}{Apple}}}}}  & 
\multicolumn{1}{c|}{\multirow{2}{*}{High Price}} & \$500,000 & & \$100,000 & \\ 
&  &\multicolumn{1}{|c}{} & \$500,000 & & \$750,000\\ \cline{3-6}
&
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Low Price}}} & \textcolor{red}{\$750,000} &  & \textcolor{red}{\$250,000} & \\ 
& & \multicolumn{1}{|c}{} & \$100,000 &  & \$250,000 \\ \cline{3-6}
\end{tabular}
\caption*{Payoffs represent profits}  
\end{figure}

### Oligopoly as Simple Prisoners' Dilemma III 

- Low Price is a dominant strategy for \textcolor{blue}{Google}

\begin{figure}
\centering 
\begin{tabular}{lrcc|cc|}
& & \multicolumn{4}{c} {\textbf{\textcolor{blue}{Google}}} \\
& & \multicolumn{2}{c}{High Price} & \multicolumn{2}{c}{\textcolor{blue}{Low Price}} \\ \cline{3-6}
\parbox[t]{2mm}{\multirow{4}{*}{\rotatebox[origin=c]{90}{\textbf{\textcolor{red}{Apple}}}}}  & 
\multicolumn{1}{c|}{\multirow{2}{*}{High Price}} & \$500,000 & & \$100,000 & \\ 
&  &\multicolumn{1}{|c}{} & \$500,000 & & \textcolor{blue}{\$750,000}\\ \cline{3-6}
&
\multicolumn{1}{c|}{\multirow{2}{*}{Low Price}} & \$750,000 &  & \$250,000 & \\ 
& & \multicolumn{1}{|c}{} & \$100,000 &  & \textcolor{blue}{\$250,000} \\ \cline{3-6}
\end{tabular}
\caption*{Payoffs represent profits}  
\end{figure}

### Oligopoly as Simple Prisoners' Dilemma IV

- Nash Equilibrium: (\textcolor{red}{Low}, \textcolor{blue}{Low})
- Pareto Improvement (for firms, not for consumers!): (\textcolor{red}{High}, \textcolor{blue}{High})
- But not a Nash Equilibrium

\begin{figure}
\centering 
\begin{tabular}{lrcc|cc|}
& & \multicolumn{4}{c} {\textbf{\textcolor{blue}{Google}}} \\
& & \multicolumn{2}{c}{\textcolor<2>{blue}{High Price}} & \multicolumn{2}{c}{\textcolor<1>{blue}{Low Price}} \\ \cline{3-6}
\parbox[t]{2mm}{\multirow{4}{*}{\rotatebox[origin=c]{90}{\textbf{\textcolor{red}{Apple}}}}}  & 
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor<2>{red}{High Price}}} & \textcolor<2>{red}{\$500,000} & & \$100,000 & \\ 
&  &\multicolumn{1}{|c}{} & \textcolor<2>{blue}{\$500,000} & & \$750,000\\ \cline{3-6}
&
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor<1>{red}{Low Price}}} & \$750,000 &  & \textcolor<1>{red}{\$250,000} & \\ 
& & \multicolumn{1}{|c}{} & \$100,000 &  & \textcolor<1>{blue}{\$250,000} \\ \cline{3-6}
\end{tabular}
\caption*{Payoffs represent profits}  
\end{figure}

### Oligopoly as Simple Prisoners' Dilemma V

\columnsbegin

\column{.5\textwidth}

- Google and Apple could **collude** with one another and agree to both raise prices
- \alert{Cartel}: group of sellers coordinate to raise prices to act like a collective monopoly and split the profits	

\column{.5\textwidth}

\begin{figure}
		\includegraphics[height=1.75in]{backroomdeal}	
\end{figure}

\columnsend

## Cartels

### Cartels

\columnsbegin

\column{.7\textwidth}

People of the same trade seldom meet together, even for merriment and diversion, but the conversation ends in a conspiracy against the public, or in some contrivance to raise prices.  \\ \medskip \emph{Wealth of Nations} (1776): Book I, Chapter 2.2

\column{.23\textwidth}

\begin{tabular}{c}
	\includegraphics[height=2in]{smith}\\
				\small Adam Smith \\
				\small (1723-1790) \\
\end{tabular}

\columnsend

### Cartels II

\columnsbegin

\column{.6\textwidth}

- Cartels often unstable:
- Incentive for each member to cheat is too strong
- Entrants (non-cartel members) can threaten lower prices
- Difficult to monitor whether firms are upholding agreement
- Cartels are illegal, must be discrete 

\column{.4\textwidth}

\begin{figure}
			\includegraphics[height=1.75in]{businesslie}
\end{figure}

\columnsend

### Sherman Antitrust Act (1890)

> Section 1: "Every contract, combination in the form of trust or otherwise, or conspiracy, in restraint of trade or commerce among the several States, or with foreign nations, is declared to be illegal."

> Section 2: "Every person who shall monopolize, or attempt to monopolize, or combine or conspire with any other person or persons, to monopolize any part of the trade or commerce among the several States, or with foreign nations, shall be deemed guilty of a felony [...]"

26 Stat. 209, 15 U.S.C. \S 1–7	


### Trying to Maintain a Cartel

\begin{figure}
		\includegraphics[height=2in]{lysinecartel}
		\includegraphics[height=2in]{theinformant}
	\caption*{Archer Daniels Midland (USA), Ajinomoto (Japan), Koywa Hakko Kogyo (Japan), Sewon American Inc (South Korea) held secret meetings to fix the price of lysine, a food additive to animal feed in the 1990s. An FBI informant brought the cartel down.}
	\end{figure}

### Trying to Maintain a Cartel II 

\begin{figure}
		\includegraphics[height=2.75in]{nflteams}
	\caption*{}
\end{figure}

### Trying to Maintain a Cartel: Coordination Devices 

\columnsbegin

\column{.5\textwidth}

- Market for turbines (for electric utility companies) in 1950s
- Three firms dominated market
    - GE: 60\%
    - Westinghouse: 30\%
    - Allied-Chalmers: 10\%
- Maintained market shares with clever coordination device

\column{.5\textwidth}

\begin{figure}
  \includegraphics[height=2in]{turbine}
\end{figure}

\columnsend

Dixit and Nalebuff (2008, 92) 

### Trying to Maintain a Cartel: Coordination Devices II 

\columnsbegin

\column{.5\textwidth}

- Utility companies put in bids to buy turbines
- If they put bid in first 1-17 days of lunar calendar:
    - Westinghouse and AC would bid prohibitively high
    - Ensure GE won
- If they put the bid in days 18-25:
    - GE and AC would bid prohibitively high
    - Ensure Westinghouse won
- If they put bid in days 26-28:
    - Ensure AC won 

\column{.5\textwidth}

\begin{figure}
  \includegraphics[height=2in]{turbine}
\end{figure}

\columnsend

Dixit and Nalebuff (2008, 92) 

### Trying to Maintain a Cartel: Coordination Devices III

\columnsbegin

\column{.5\textwidth}

- Utility companies did *not* release bids according to lunar calendar...ensured distribution
- Cheating by one firm easily discoverable by others
- DOJ didn't think to link to lunar calendar...until they did

\column{.5\textwidth}

\begin{figure}
  \includegraphics[height=2in]{turbine}
\end{figure}

\columnsend

Dixit and Nalebuff (2008, 92) 

### Trying to Maintain a Cartel: Coordination Devices IV

\columnsbegin

\column{.5\textwidth}

- Bidding for electromagnetic spectrum licenses 1996-1997
- Firm that wanted license in a particular location would signal to other firms its determination to fight for that location by including the telephone area code in the last three digits of its bid
    - e.g. \$50,100,**202** for Washington D.C. (area code **202**)
- Other firms would let in win (in exchange for them doing likewise)

\column{.5\textwidth}

\begin{figure}
  \includegraphics[height=2in]{signal}
\end{figure}

\columnsend

Dixit and Nalebuff (2008, 92-93) 

### Trying to Maintain a Cartel: Price Matching

\columnsbegin

\column{.5\textwidth}

- One of the hardest parts to enforce a cartel is to detect cheating or chiseling by members lowering prices
- Detection and retaliation of retail price-cutting made automatic with "matching or beating the competition" policies or "most favored customer" clauses
- Loudly claim to match or beat any competitors' price
- Appears to promote competition by guaranteeing low prices!
- But may actually have the opposite effect if the stores are secretly a cartel

\column{.5\textwidth}

\begin{figure}
  \includegraphics[height=1.5in]{priceguarantee}\\
  \includegraphics[height=1.5in]{pricematch}
\end{figure}

\columnsend

Dixit and Nalebuff (2008, 93-94) 

### Trying to Maintain a Cartel: Price Matching II 

\columnsbegin

\column{.5\textwidth}

- Suppose L.L. Bean and Land's End tacitly agree to sell flannel shirts above market price at \$80
- Both have "match or beat the competition" policies
- If one firm quickly sneaks a \$70 sale on the shirt, the other will find out quickly and retaliate
- Extra clever: use customer, who has incentive to find lowest price, to detect cheating! 

\column{.5\textwidth}

\begin{figure}
  \includegraphics[height=1.5in]{priceguarantee}\\
  \includegraphics[height=1.5in]{pricematch}
\end{figure}

\columnsend

Dixit and Nalebuff (2008, 93-94) 

### Cartels: Support by Government

- Some cartels persist *because* they are tacitly or explicitly supported by governments

\begin{figure}
		\includegraphics[height=1.5in]{nra}
		\caption*{National Industrial Recovery Act of 1933 cartelized most major industries to artificially raise prices of goods. It was ultimately found unconstitutional (in \emph{Schechter Poultry Corp. v. United States} (1935))} 	
	\end{figure}

### Cartels: Support by Government II 

\begin{figure}
		\includegraphics[height=1.5in]{marvinhorne}
		\caption*{Marvin Horne, a small raisin grower, recently won a Supreme Court case allowing him to refuse handing over his raisins to the Raisin Administrative Committee (yes, that's a thing) which confiscates raisins from farmers and holds them off from the market to raise raisin prices under ``the world's most outdated law'' (\href{http://www.forbes.com/sites/trevorburrus/2015/04/24/rebel-farmers-government-cartels-how-the-new-deal-cartelized-u-s-agriculture/}{Forbes}, \href{http://www.npr.org/sections/money/2015/06/24/417187957/episode-478-the-raisin-outlaw}{NPR})} 
	\end{figure}


### Cartels: Support by Government III 

\begin{figure}
		\includegraphics[height=1.5in]{icc}
		\includegraphics[height=1.5in]{robberbarons}
		\caption*{The Interstate Commerce Commission (ICC) was set up to regulate the ``robber barons''' monopolistic railroads. But economic historians widely believe it actually cartelized a fiercely competitive industry and helped it keep rates artificially high!} 
	\end{figure}

### Cartels: Support by Government IV

\begin{columns}
		\begin{column}[c]{7cm}
			\small ``[B]ecause of their inability to maintain their cartels [prior to the ICC], railroads were big supporters of the [Interstate Commerce Act] because the newly-formed ICC could coordinate cartel prices...Using the new law as authority, the railroads revamped their freight classification, raised rates, eliminated passes and fare reductions, and revised less than carload rates on all types of goods, including groceries.''
			\\ \medskip (1963). \emph{The Triumph of Conservatism: A Reinterpretation of American History, 1900-1916}
		\end{column}
		\begin{column}[c]{3cm}
			\begin{tabular}{c}
				\includegraphics[height=1.75in]{kolko}\\
			\small Gabriel Kolko \\
			\small (1932-2014)\\
			\end{tabular}
		\end{column}	
\end{columns}

### Cartels: Support by Government V

\begin{figure}
		\includegraphics[height=1.5in]{opeccountries}
		\includegraphics[height=1.5in]{opecflag}
	\end{figure}

### Cartels: Support by Government VI

\begin{figure}
		\includegraphics[height=1.75in]{tuco}
		\includegraphics[height=1.75in]{drugcartels}
		\caption*{} 	
\end{figure}

### Cartels: Support by Government VII

\columnsbegin

\column{.65\textwidth}

"See, if you look at the drug war from a purely economic point of view, the role of the government is to protect the drug cartel. That's literally true."

\column{.35\textwidth}

\begin{tabular}{c}
  \includegraphics[height=2in]{friedman}\\
  Milton Friedman\\
  (1912--2006)\\
  Economics Nobel 1976\\
\end{tabular}

\columnsend 

[1991 Interview](https://www.ukcia.org/research/argue/milton.htm)

### Cartels: Support by Government VIII

\columnsbegin

\column{.65\textwidth} 

\scriptsize

>[I]t’s very hard for a small person to go into the drug importing business because our interdiction efforts essentially make it enormously costly. So, the only people who can survive in that business are these large Medellin cartel kind of people who have enough money so they can have fleets of airplanes, so they can have sophisticated methods, and so on.

>In addition to which, by keeping goods out and by arresting, let's say, local marijuana growers, the government keeps the price of these products high. What more could a monopolist want? He's got a government who makes it very hard for all his competitors and who keeps the price of his products high. It's absolutely heaven."

\column{.35\textwidth}

\begin{tabular}{c}
  \includegraphics[height=2in]{friedman}\\
  Milton Friedman\\
  (1912--2006)\\
  Economics Nobel 1976\\
\end{tabular}

\columnsend 

[1991 Interview](https://www.ukcia.org/research/argue/milton.htm)

### Cartels in Fiction

\begin{figure}
		\includegraphics[height=2.25in]{thewirestringer}
		\caption*{\href{https://www.youtube.com/watch?v=hGo5bxWy21g}{\textcolor{magenta}{The Wire: The New Day Co-Op}}}
	\end{figure}

### Cartels in Fiction II 

\begin{figure}
		\includegraphics[height=2.25in]{thewirestringer}
		\caption*{\href{https://www.youtube.com/watch?v=aT7TxMaZ4eM}{\textcolor{magenta}{The Wire: Marlo Dismantles the Co-Op}}}
	\end{figure}

## Bertrand Competition

### Better Theories of Oligopoly

1. Bertrand competition: firms *simultaneously* compete on *price*
2. Cournot competition: firms *simultaneously* compete on *quantity*
3. Stackelberg competition: firms *sequentially* compete on *quantity*

### Bertrand Competition

\columnsbegin

\column{.75\textwidth}

- **"Bertrand competition"**: two (or more) firms compete on *price* to sell the same good 
- Firms set their prices simultaneously
- Consumers are indifferent between the brands and always buy from the seller with the lowest price
- Bertrand actually responding to a more famous theory by Cournot (we'll see next)
    - We start with Bertrand to ease our way into these models (Bertrand is easier)

\column{.25\textwidth}

\begin{tabular}{c}
\includegraphics[height=2in]{Bertrand}\\
Joseph Bertrand\\
(1822--1990)\\
France\\
\end{tabular}

\columnsend

### Bertrand Competition I

\begin{example}
\end{example}

- Suppose two firms, \textcolor{red}{ } and \textcolor{blue}{Target} sell identical HDTVs
- Suppose it costs each firm \$200 to stock an HDTV
- Let $Q$ be the *total* quantity purchased by consumers from the entire market (i.e. both firms)  
- Denote \textcolor{red}{Walmart}'s price as \textcolor{red}{$p_w$} and \textcolor{blue}{Target}'s price as \textcolor{blue}{$p_t$}

\columnsbegin

\column{.5\textwidth}

- Demand for HDTVs at \textcolor{red}{Walmart}
- $Q$ if $\textcolor{red}{p_w}<\textcolor{blue}{p_t}$
- $\frac{Q}{2}$ if $\textcolor{red}{p_w}=\textcolor{blue}{p_t}$
- 0 if $\textcolor{red}{p_w}>\textcolor{blue}{p_t}$

\column{.5\textwidth}

- Demand for HDTVs at \textcolor{blue}{Target} 
- $Q$ if $\textcolor{red}{p_w}>\textcolor{blue}{p_t}$
- $\frac{Q}{2}$ if $\textcolor{red}{p_w}=\textcolor{blue}{p_t}$
- 0 if $\textcolor{red}{p_w}<\textcolor{blue}{p_t}$

\columnsend

- The only way to sell HDTVs is to match or beat your competitor

### Bertrand Competition II

\columnsbegin

\column{.5\textwidth}

- If you are \textcolor{red}{Walmart}, setting your price $\textcolor{red}{p_w}=\textcolor{blue}{p_t}-\epsilon$ for any arbitrarily small $\epsilon>0$ will capture you the entire market
- Same for \textcolor{blue}{Target} vs. you
- Won't charge $p<MC$, since that will lose profits
- The only price that cannot be undercut is setting $p_w=p_t=MC$! 
- **Nash Equilibrium**: all firms charge (identical) MC (and earn no profits)
    - This is the perfectly competitive outcome, but with just 2 firms!

\column{.5\textwidth}

\begin{figure}
  \includegraphics[height=1.5in]{priceguarantee}\\
  \includegraphics[height=1.5in]{pricematch}
\end{figure}

\columnsend

### Bertrand Competition: Reaction Curves

\columnsbegin

\column{.4\textwidth}

\scriptsize 

- We can graph \textcolor{red}{Walmart}'s **reaction curve** to \textcolor{blue}{Target}'s price
- e.g. if \textcolor{blue}{Target} sets a price of \textcolor{blue}{\$500}, \textcolor{red}{Walmart}'s best response is \textcolor{red}{$\$500-\epsilon$}
- e.g. if \textcolor{blue}{Target} sets a price of \textcolor{blue}{\$300}, \textcolor{red}{Walmart}'s best response is \textcolor{red}{$\$300-\epsilon$}
- e.g. if \textcolor{blue}{Target} sets a price of \textcolor{blue}{\$200} (MC), \textcolor{red}{Walmart}'s best response is \textcolor{red}{$\$200$} (MC) 

\column{.6\textwidth}

\begin{figure}
\centering
\begin{tikzpicture}\scriptsize 
	\begin{axis}[
		scale=0.9,
		axis lines=middle, 
		enlarge x limits={rel=0.1, upper},
		enlarge y limits={rel=0.1, upper},
		every axis y label/.style={at={(axis description cs:-0.2,0.5)},rotate=90,anchor=north},
		every axis x label/.style={at={(axis description cs:0.5,-0.1)},anchor=north},
	legend style={at={(0.5,-0.25)},anchor=north},
	xlabel=\textcolor{blue}{Target's Price ($p_t$)},
	ylabel=\textcolor{red}{Walmart's Price ($p_w$)},
	xtick={100, 200,...,1000},
	ytick={100, 200,...,1000},
	grid=major,
	ymin=0,
	xmin=0,
	ymax=1000,
	xmax=1000,
	legend cell align=left,
]
	\addplot[domain=0:1000, thick, dotted]{x};
	\addlegendentry{$45^\circ$}
	\draw[ultra thick, red](axis cs:0,200)--(axis cs:225,200)--(axis cs:1025,1000);
	\addlegendimage{ultra thick, red}
	\addlegendentry{Walmart's Reaction Curve}
	\draw<2->[very thick, dashed, blue] (axis cs:500,0)--(axis cs:500,475);
	\draw<2->[very thick, dashed, red] (axis cs:500,475)--(axis cs:0,475);
	\draw<3->[very thick, dashed, blue] (axis cs:300,0)--(axis cs:300,275);
	\draw<3->[very thick, dashed, red] (axis cs:300,275)--(axis cs:0,275);
	\draw<4->[very thick, dashed, blue] (axis cs:200,0)--(axis cs:200,200);
\end{axis}
\end{tikzpicture}	
\end{figure}

\columnsend

### Bertrand Competition: Reaction Curves II

\columnsbegin

\column{.4\textwidth}

\scriptsize 

- Look also at \textcolor{blue}{Target}'s firm's **reaction curve** to \textcolor{red}{Walmart}'s price
- e.g. if \textcolor{red}{Walmart} sets a price of \textcolor{red}{\$500}, \textcolor{blue}{Target}'s best response is \textcolor{blue}{$\$500-\epsilon$}
- e.g. if \textcolor{blue}{Walmart} sets a price of \textcolor{red}{\$300}, \textcolor{blue}{Target}'s best response is \textcolor{blue}{$\$300-\epsilon$}
- e.g. if \textcolor{blue}{Walmart} sets a price of \textcolor{blue}{\$200} (MC), \textcolor{blue}{Target}'s best response is \textcolor{red}{$\$200$} (MC) 

\column{.6\textwidth}

\begin{figure}
\centering
\begin{tikzpicture}\scriptsize 
	\begin{axis}[
		scale=0.9,
		axis lines=middle, 
		enlarge x limits={rel=0.1, upper},
		enlarge y limits={rel=0.1, upper},
		every axis y label/.style={at={(axis description cs:-0.2,0.5)},rotate=90,anchor=north},
		every axis x label/.style={at={(axis description cs:0.5,-0.1)},anchor=north},
	legend style={at={(0.5,-0.25)},anchor=north},
	xlabel=\textcolor{blue}{Target's Price ($p_t$)},
	ylabel=\textcolor{red}{Walmart's Price ($p_w$)},
	xtick={100, 200,...,1000},
	ytick={100, 200,...,1000},
	grid=major,
	ymin=0,
	xmin=0,
	ymax=1000,
	xmax=1000,
	legend cell align=left,
]
	\addplot[domain=0:1000, thick, dotted]{x};
	\addlegendentry{$45^\circ$}
\draw[ultra thick, blue](axis cs:200,0)--(axis cs:200,225)--(axis cs:1000,1025);
	\addlegendimage{ultra thick, blue}
	\addlegendentry{Target's Reaction Curve}
	\draw<2->[very thick, dashed, red] (axis cs:0,500)--(axis cs:475,500);
	\draw<2->[very thick, dashed, blue] (axis cs:475,500)--(axis cs:475,0);
	\draw<3->[very thick, dashed, red] (axis cs:0,300)--(axis cs:275,300);
	\draw<3->[very thick, dashed, blue] (axis cs:275,300)--(axis cs:275,0);
	\draw<4->[very thick, dashed, red] (axis cs:0,200)--(axis cs:200,200);
\end{axis}
\end{tikzpicture}	
\end{figure}

\columnsend

### Bertrand Competition: Reaction Curves II

\columnsbegin

\column{.4\textwidth}

- Combine both curves on the same graph
- **Nash equilibrium**: (\textcolor{red}{\$200}, \textcolor{blue}{\$200}) i.e. (\textcolor{red}{MC}, \textcolor{blue}{MC})
    - Where both reaction curves intersect
- No longer an incentive to undercut or change price 

\column{.6\textwidth}

\begin{figure}
\centering
\begin{tikzpicture}\scriptsize 
	\begin{axis}[
		scale=0.9,
		axis lines=middle, 
		enlarge x limits={rel=0.1, upper},
		enlarge y limits={rel=0.1, upper},
		every axis y label/.style={at={(axis description cs:-0.2,0.5)},rotate=90,anchor=north},
		every axis x label/.style={at={(axis description cs:0.5,-0.1)},anchor=north},
	legend style={at={(0.5,-0.25)},anchor=north},
	xlabel=\textcolor{blue}{Target's Price ($p_t$)},
	ylabel=\textcolor{red}{Walmart's Price ($p_w$)},
	xtick={100, 200,...,1000},
	ytick={100, 200,...,1000},
	grid=major,
	ymin=0,
	xmin=0,
	ymax=1000,
	xmax=1000,
	legend cell align=left,
]
	\addplot[domain=0:1000, thick, dotted]{x};
	\addlegendentry{$45^\circ$}
	\draw[ultra thick, red](axis cs:0,200)--(axis cs:225,200)--(axis cs:1025,1000);
	\addlegendimage{ultra thick, red}
	\addlegendentry{Walmart's Reaction Curve}
	\draw[ultra thick, blue](axis cs:200,0)--(axis cs:200,225)--(axis cs:1000,1025);
	\addlegendimage{ultra thick, blue}
	\addlegendentry{Target's Reaction Curve}
	\draw<2->[fill=black] (axis cs:200,200)circle(0.075cm)node[right]{$N.E.$};
\end{axis}
\end{tikzpicture}	
\end{figure}

\columnsend

### Bertrand Competition Game (Moblab)

\begin{figure}
  \includegraphics[height=2.75in]{Bertrandmoblab}
\end{figure}

### Bertrand Competition Game (Moblab) II

\columnsbegin

\column{.7\textwidth}

- Each of you is a firm selling identical Economics course notes
- You will be put into a market with one other player (firm)
- Each term, both of you simultaneously choose your price
- The lowest price $p_L$ determines the market demand
$$q_d=3600-200p_L$$
- Firm(s) choosing the lowest price get **all** the customers
- Both firms have \$2 cost per unit sold
- $p=\$10$ maximizes total market profits

\column{.3\textwidth}

\begin{figure}
  \includegraphics[height=1.25in]{Bertrandmoblab}
\end{figure}

\columnsend

### Bertrand Competition Game (Moblab) III

\columnsbegin

\column{.7\textwidth}

\begin{example}
\end{example}

$$q_d=3600-200p_L$$

- Suppose Firm 1 sets $p=\$9$ and Firm 2 sets $p=\$10$
- Firm 2 sells 0, makes \$0
- Firm 1 sells $q_d=3,600-200(9)=1,800$ and earns $1,800(\$9-\$2)=\$12,600$

\column{.3\textwidth}

\begin{figure}
  \includegraphics[height=1.25in]{Bertrandmoblab}
\end{figure}

\columnsend

## Cournot Competition

### Cournot Competition 

\columnsbegin

\column{.75\textwidth}

- **"Cournot competition"**: two (or more) firms compete on *quantity* to sell the same good 
- Firms set their *quantities* simultaneously
- Firms' joint output determines the market price faced by all firms 

\column{.25\textwidth}

\begin{tabular}{c}
\includegraphics[height=2in]{Cournot}\\
Antoine Augustin Cournot\\
(1801--1877)\\
France\\
\end{tabular}

\columnsend

### Cournot Competition II

\begin{example}

\end{example}

- Suppose two firms, each have an identical constant $MC=AC=c$
- Firms 1 and 2 simultaneously choose production quantities $q_1$ and $q_2$
- The total market (inverse) demand is given by:
\vspace{-0.5cm}\begin{align*}
P&=a-bQ\\
Q&=q_1+q_2\\
\end{align*}\vspace{-0.5cm}

- A firm (e.g. firm 1)'s profit is found by: 
\vspace{-0.5cm}\begin{align*}
\pi_1&=q_1(P-c)\\
\pi_1&=q_1(a-b(q_1+q_2)-c)\\
\end{align*}\vspace{-0.5cm}
- Note each firm's profits depend (in part) on the outputs of the other firm! 

### Cournot Competition III

\begin{example}
\end{example}

- Assume Saudi Arabia ($sa$) and Iran ($i$) are the only two oil producers, each with a constant $MC=AC=\$20$. The market (inverse) demand curve is given by:
\vspace{-0.5cm}\begin{align*}
P&=200-3Q\\
Q&=q_{sa}+q_i\\
\only<2->{P&=200-3q_{sa}-3q_i\\}
\end{align*}\vspace{-0.5cm}
\onslide<3->
- Firms maximize profit by setting $q^*: MR=MC$
- Solving for Saudi Arabia, recall that MR is twice the slope of the inverse demand curve...
$$MR_{sa}=200-6q_{sa}-3q_i$$
    - Saudi Arabia considers Iran's output as a constant, not their own choice variable
    - This is **"residual demand"** for Saudi Arabian oil (and finding $MR_{sa}$) after *subtracting out* Iran's $q_i^*$

### Cournot Competition IV

- By solving for $q^*$ for each firm, where $MR=MC$, we find each firm's **reaction function** or **best response function**
- With identical marginal costs, they are symmetric: 
\vspace{-0.5cm}\begin{align*}
q_{sa}^*&=30-0.5q_i\\
q_i^*&=30-0.5q_{sa}\\
\end{align*}\vspace{-0.5cm}

### Cournot Competition V

\columnsbegin

\column{.4\textwidth}

- We can graph \textcolor{red}{Saudi Arabia}'s **reaction curve** to \textcolor{blue}{Iran}'s output
- e.g. if \textcolor{blue}{Iran} produces \textcolor{blue}{40}, \textcolor{red}{Saudi Arabias}'s best response is \textcolor{red}{10}
- e.g. if \textcolor{blue}{Iran} produces \textcolor{blue}{20}, \textcolor{red}{Saudi Arabias}'s best response is \textcolor{red}{20}

\column{.6\textwidth}

\begin{figure}
\centering
\begin{tikzpicture}\scriptsize 
	\begin{axis}[
		scale=0.9,
		axis lines=middle, 
		enlarge x limits={rel=0.1, upper},
		enlarge y limits={rel=0.1, upper},
		every axis y label/.style={at={(axis description cs:-0.2,0.5)},rotate=90,anchor=north},
		every axis x label/.style={at={(axis description cs:0.5,-0.1)},anchor=north},
	legend style={at={(0.5,-0.25)},anchor=north},
	xlabel=\textcolor{blue}{Iran's output ($q_{i}$)},
	ylabel=\textcolor{red}{Saudi Arabia's output ($q_{sa}$)},
	xtick={0,5,...,60},
	ytick={0,5,...,60},
	grid=major,
	ymin=0,
	xmin=0,
	ymax=60,
	xmax=60,
	legend cell align=left,
]
	\addplot[domain=0:60, ultra thick, red]{30-0.5*x};
	\addlegendentry{Saudi Arabia's Reaction Curve}
	\draw<3->[very thick, dotted, red](axis cs:0,20)--(axis cs:20,20);
	\draw<3->[very thick, dotted, blue](axis cs:20,20)--(axis cs:20,0);
	\draw<2->[very thick, dotted, blue](axis cs:40,0)--(axis cs:40,10);
	\draw<2->[very thick, dotted, red](axis cs:40,10)--(axis cs:0,10);
	\end{axis}
\end{tikzpicture}
\end{figure}

\columnsend

### Cournot Competition VI

\columnsbegin

\column{.4\textwidth}

- Look also at \textcolor{blue}{Iran}'s **reaction curve** to \textcolor{red}{Saudi Arabia}'s output
- e.g. if \textcolor{red}{Saudi Arabia} produces \textcolor{red}{40}, \textcolor{blue}{Iran}'s best response is \textcolor{blue}{10}
- e.g. if \textcolor{red}{Saudi Arabia} produces \textcolor{red}{20}, \textcolor{blue}{Iran}'s best response is \textcolor{blue}{20}

\column{.6\textwidth}

\begin{figure}
\centering
\begin{tikzpicture}\scriptsize 
	\begin{axis}[
		scale=0.9,
		axis lines=middle, 
		enlarge x limits={rel=0.1, upper},
		enlarge y limits={rel=0.1, upper},
		every axis y label/.style={at={(axis description cs:-0.2,0.5)},rotate=90,anchor=north},
		every axis x label/.style={at={(axis description cs:0.5,-0.1)},anchor=north},
	legend style={at={(0.5,-0.25)},anchor=north},
	xlabel=\textcolor{blue}{Iran's output ($q_{i}$)},
	ylabel=\textcolor{red}{Saudi Arabia's output ($q_{sa}$)},
	xtick={0,5,...,60},
	ytick={0,5,...,60},
	grid=major,
	ymin=0,
	xmin=0,
	ymax=60,
	xmax=60,
	legend cell align=left,
]
	\addplot[domain=0:60, ultra thick, blue]{60-2*x};
	\addlegendentry{Iran's Reaction Curve}
	\draw<2->[very thick, dotted, red](axis cs:0,40)--(axis cs:10,40);
	\draw<2->[very thick, dotted, blue](axis cs:10,40)--(axis cs:10,0);
	\draw<3->[very thick, dotted, blue](axis cs:20,0)--(axis cs:20,20);
	\draw<3->[very thick, dotted, red](axis cs:20,20)--(axis cs:0,20);
	\end{axis}
\end{tikzpicture}
\end{figure}

\columnsend

### Cournot Competition VII

\columnsbegin

\column{.4\textwidth}

- Put them together
- **Nash equilibrium** where both reaction curves intersect: (\textcolor{red}{20}, \textcolor{blue}{20})
- Algebraically: plug one firm's reaction curve into the other's

\column{.6\textwidth}

\begin{figure}
\centering
\begin{tikzpicture}\scriptsize 
	\begin{axis}[
		scale=0.9,
		axis lines=middle, 
		enlarge x limits={rel=0.1, upper},
		enlarge y limits={rel=0.1, upper},
		every axis y label/.style={at={(axis description cs:-0.2,0.5)},rotate=90,anchor=north},
		every axis x label/.style={at={(axis description cs:0.5,-0.1)},anchor=north},
	legend style={at={(0.5,-0.25)},anchor=north},
	xlabel=\textcolor{blue}{Iran's output ($q_{i}$)},
	ylabel=\textcolor{red}{Saudi Arabia's output ($q_{sa}$)},
	xtick={0,5,...,60},
	ytick={0,5,...,60},
	grid=major,
	ymin=0,
	xmin=0,
	ymax=60,
	xmax=60,
	legend cell align=left,
]
	\addplot[domain=0:60, ultra thick, red]{30-0.5*x};
	\addlegendentry{Saudi Arabia's Reaction Curve}
	\addplot[domain=0:60, ultra thick, blue]{60-2*x};
	\addlegendentry{Iran's Reaction Curve}
	\draw[fill=black] (axis cs:20,20)circle(0.075cm)node[right]{$N.E.$};
	\draw[very thick, dotted](axis cs:0,20)--(axis cs:20,20)--(axis cs:20,0);
	\end{axis}
\end{tikzpicture}
\end{figure}

\columnsend

### Cournot Competition VIII

- The market price of oil again was
$$P=200-3q_{sa}-3q_i$$
- Both countries produce 20:
\vspace{-0.5cm}\begin{align*}
P&=200-3(20)-3(20)\\
P&=\$80\\
\end{align*}\vspace{-0.5cm}
- We can find the profit for each country
\vspace{-0.5cm}\begin{align*}
\pi_{sa}&=q_{sa}(P-c)\\
\pi_{sa}&=20(80-20)\\
\pi_{sa}&=\$1,200\\
\end{align*}\vspace{-0.5cm}
- Symmetrically, for Iran, $\pi_i=\$1,200$

### Cournot Collusion

- Suppose now both firms **collude** and act like a single monopolist, who sets:
\vspace{-0.6cm}\begin{align*}
MR&=MC\\
200-6Q&=20\\
30&=Q^*\\
\end{align*}\vspace{-0.6cm}
- The monopoly price will then be
\vspace{-0.6cm}\begin{align*}
P&=200-3(30)\\
P&=\$110\\
\end{align*}\vspace{-0.6cm}
- Total profit will then be:
\vspace{-0.6cm}\begin{align*}
\Pi&=30(110-20)\\
\Pi&=\$2,700\\
\end{align*}\vspace{-0.6cm}
with \$1,400 going to each firm

### Cournot Collusion II

\columnsbegin

\column{.4\textwidth}

- **Cournot Competition**: each firm produces 20 and earns \$1,200
- **Collusion**: each firm produces 15 and earns \$1,400 

\column{.6\textwidth}

\begin{figure}
\centering
\begin{tikzpicture}\scriptsize 
	\begin{axis}[
		scale=0.9,
		axis lines=middle, 
		enlarge x limits={rel=0.1, upper},
		enlarge y limits={rel=0.1, upper},
		every axis y label/.style={at={(axis description cs:-0.2,0.5)},rotate=90,anchor=north},
		every axis x label/.style={at={(axis description cs:0.5,-0.1)},anchor=north},
	legend style={at={(0.5,-0.25)},anchor=north},
	xlabel=\textcolor{blue}{Iran's output ($q_{i}$)},
	ylabel=\textcolor{red}{Saudi Arabia's output ($q_{sa}$)},
	xtick={0,5,...,60},
	ytick={0,5,...,60},
	grid=major,
	ymin=0,
	xmin=0,
	ymax=60,
	xmax=60,
	legend cell align=left,
]
	\addplot[domain=0:60, ultra thick, red]{30-0.5*x};
	\addlegendentry{Saudi Arabia's Reaction Curve}
	\addplot[domain=0:60, ultra thick, blue]{60-2*x};
	\addlegendentry{Iran's Reaction Curve}
	\draw[fill=black] (axis cs:20,20)circle(0.075cm)node[right]{$N.E.$};
	\draw[very thick, dotted](axis cs:0,20)--(axis cs:20,20)--(axis cs:20,0);
	\draw[fill=black] (axis cs:15,15)circle(0.075cm)node[above]{$Collusion$};
	\draw[very thick, dotted](axis cs:0,15)--(axis cs:15,15)--(axis cs:15,0);
	\end{axis}
\end{tikzpicture}
\end{figure}

\columnsend

### Cournot Collusion II

\columnsbegin

\column{.4\textwidth}

- But is collusion a Nash Equilibrium?
- Suppose Saudi Arabia knows Iran is producing 15 (for collusion)
- Saudi Arabia would then want to produce 22.5

\column{.6\textwidth}

\begin{figure}
\centering
\begin{tikzpicture}\scriptsize 
	\begin{axis}[
		scale=0.9,
		axis lines=middle, 
		enlarge x limits={rel=0.1, upper},
		enlarge y limits={rel=0.1, upper},
		every axis y label/.style={at={(axis description cs:-0.2,0.5)},rotate=90,anchor=north},
		every axis x label/.style={at={(axis description cs:0.5,-0.1)},anchor=north},
	legend style={at={(0.5,-0.25)},anchor=north},
	xlabel=\textcolor{blue}{Iran's output ($q_{i}$)},
	ylabel=\textcolor{red}{Saudi Arabia's output ($q_{sa}$)},
	xtick={0,5,...,60},
	ytick={0,5,...,60},
	grid=major,
	ymin=0,
	xmin=0,
	ymax=60,
	xmax=60,
	legend cell align=left,
]
	\addplot[domain=0:60, ultra thick, red]{30-0.5*x};
	\addlegendentry{Saudi Arabia's Reaction Curve}
	\addplot[domain=0:60, ultra thick, blue]{60-2*x};
	\addlegendentry{Iran's Reaction Curve}
	\draw[fill=black] (axis cs:20,20)circle(0.075cm)node[right]{$N.E.$};
	\draw[fill=black] (axis cs:15,15)circle(0.075cm)node[above]{$Collusion$};
	\draw<2->[ultra thick, dashed, blue](axis cs:15,0)--(axis cs:15,15);
	\draw<3->[ultra thick, dashed, blue](axis cs:15,15)--(axis cs:15,22.5);
	\draw<3->[ultra thick, dashed, red](axis cs:15,22.5)--(axis cs:0,22.5);
	\end{axis}
\end{tikzpicture}
\end{figure}

\columnsend

### Cournot Collusion III

- This would bring market price to
\vspace{-0.5cm}\begin{align*}
P&=200-3q_{sa}-3q_i\\
P&=200-3(22.5)-3(15)\\
P&=\$87.50\\
\end{align*}\vspace{-0.5cm}

\columnsbegin

\column{.5\textwidth}

- Saudi Arabia's profit would be 
\vspace{-0.5cm}\begin{align*}
\pi_{sa}&=q_{sa}(P-c)\\
\pi_{sa}&=22.5(87.50-20)\\
\pi_{sa}&=\$1,518.75\\
\end{align*}\vspace{-0.5cm}

\column{.5\textwidth}

- Iran's profit would be
\vspace{-0.5cm}\begin{align*}
\pi_{i}&=q_{i}(P-c)\\
\pi_{i}&=15(87.50-20)\\
\pi_{i}&=\$712.50\\
\end{align*}\vspace{-0.5cm}

\columnsend

### Bertrand Competition in Our Example

- Imagine *Bertrand* competition between Saudi Arabia and Iran instead (price competition)
- **Nash equilibrium**: Firms will set $P=MC$, so:
\vspace{-0.5cm}\begin{align*}
P&=MC\\
200-3Q&=20\\
Q&=60\\
\end{align*}\vspace{-0.5cm}
- Both countries split demand equally, each selling 30 units
- Profit for both countries would be 0, since $P=MC$

### Cournot vs. Bertrand Models

\begin{table}
\begin{tabular}{lrrr}
		Oligopoly & Industry & Market & Industry\\
	 	Type & Output & Price & Profit\\ \toprule
		Collusion & 30 & \$110 & \$2,700 \\ 
		Cournot & 40 & \$80 & \$2,400\\
		Bertrand & 60 & \$20 & \$0\\ \bottomrule 	
\end{tabular}
\end{table}

- Output: $Q_m < Q_c < Q_b$
- Market price: $P_b < P_c < P_m$
- Profit: $\pi_b=0 < \pi_c < \pi_m$

Where subscript $m$ is monopoly (collusion), $c$ is Cournot, $b$ is Bertrand

### Cournot Competition, You Try

\begin{example}
Suppose Firm 1 and Firm 2 have a constant $MC=AC=\$8$. The market (inverse) demand curve is given by:
\vspace{-0.5cm}\begin{align*}
P&=200-2Q\\
Q&=q_1+q_2\\
\end{align*}\vspace{-0.5cm}

a. Find the Cournot-Nash equilibrium output and profit for each firm.

b. Find the output and profit for each firm if the two were to collude.

c. Find the price and output if the two were to compete on price instead of quantity.

\end{example}

### Cournot Competition Game (Moblab)

\begin{figure}
  \includegraphics[height=2.75in]{cournotmoblab}
\end{figure}

### Cournot Competition Game (Moblab) II

\columnsbegin

\column{.7\textwidth}

- Each of you is a firm selling identical scooters
- Each season, each firm chooses its quantity to produce
- You pay a cost for each you produce (identical across all firms)
- Market price depends on *total* industry output
    - More total output$\implies$lower market price
    - Market price is revealed after all firms have chosen their output

\column{.3\textwidth}

\begin{figure}
  \includegraphics[height=1.25in]{cournotmoblab}
\end{figure}

\columnsend

### Cournot Competition Game (Moblab) III

\columnsbegin

\column{.7\textwidth}

- We will play 4 times:
    1. You are the only firm (monopoly)
    2. You will be matched with another firm (duopoly)
    3. You will be matched with 2 other firms (triopoly)
    4. The entire class is competing in the same market ($N=22$)
- Each instance will have 3 rounds 

\column{.3\textwidth}

\begin{figure}
  \includegraphics[height=1.25in]{cournotmoblab}
\end{figure}

\columnsend

## Stackelberg Competition

### Stackelberg Competition

\columnsbegin

\column{.75\textwidth}

- Cournot-style competition: firms compete on *quantity*
- Firms' joint output determines the market price faced by all firms 
- Firms *sequentially* set their quantities:
    - **Leader** goes first
    - **Follower** goes last, must react to Leader's $q$

\column{.25\textwidth}

\begin{tabular}{c}
\includegraphics[height=2in]{stackelberg}\\
Heinrich von Stackelberg\\
(1905--1946)\\
Germany\\
\end{tabular}

\columnsend

### Stackelberg Competition: Example 

\begin{example}

\end{example}

- Continue with Saudi Arabia and Iran
- The total market (inverse) demand was given by:
$$P=200-3q_{sa}-3q_i$$
- We solved for Saudi Arabia's reaction function (and symmetrically, Iran's):
\vspace{-0.5cm}\begin{align*}
q_{sa}^*&=30-0.5q_i\\
q_i^*&=30-0.5q_{sa}\\
\end{align*}\vspace{-0.5cm}
- Suppose Saudi Arabia is the **Stackelberg leader** and sets its output first
- Saudi Arabia knows exactly how Iran will respond: $q_i^*=30-0.5q_{sa}$

### Stackelberg Competition: Example II

\begin{example}

\end{example}

- Market leader goes first, essentially faces the entire market demand
    - but not a monopolist:
    - knows that follower will still produce (and push down the market price) 
- Substitute follower's reaction function into market (inverse) demand function
\vspace{-0.5cm}\begin{align*}
P&=200-3q_{sa}-3(30-0.5q_{sa})\\
P&=110-1.5q_{sa}\\
\end{align*}\vspace{-0.5cm}
- Find $MR$ for Saudi Arabia from market demand
$$MR_{sa}=110-3q_{sa}$$

### Stackelberg Competition: Example III

\begin{example}

\end{example}

- Find $q^*: MR=MC$
\vspace{-0.5cm}\begin{align*}
MR&=MC\\
110-3q_{sa}&=20\\
30&=q_{sa}^*\\
\end{align*}\vspace{-0.5cm}
- Iran will respond
\vspace{-0.5cm}\begin{align*}
q_i^*&=30-0.5q_{sa}\\
q_i^*&=30-0.5(30)\\
q_i^*&=15\\
\end{align*}\vspace{-0.5cm}

### Stackelberg Competition: Example IV

\begin{figure}
\centering
\begin{tikzpicture}\scriptsize 
	\begin{axis}[
		scale=0.9,
		axis lines=middle, 
		enlarge x limits={rel=0.1, upper},
		enlarge y limits={rel=0.1, upper},
		every axis y label/.style={at={(axis description cs:-0.2,0.5)},rotate=90,anchor=north},
		every axis x label/.style={at={(axis description cs:0.5,-0.1)},anchor=north},
	legend style={at={(0.5,-0.25)},anchor=north},
	xlabel=Iran's output ($q_i$),
	ylabel=Saudi Arabia's output ($q_{sa}$),
	xtick={0,10,...,60},
	ytick={0,10,...,60},
	grid=major,
	ymin=0,
	xmin=0,
	ymax=60,
	xmax=60,
	legend cell align=left]
	\addplot[domain=0:60, ultra thick, red]{30-0.5*x};
	\addlegendentry{Saudi Arabia's Reaction Curve}
	\addplot[domain=0:60, ultra thick, blue]{60-2*x};
	\addlegendentry{Iran's Reaction Curve}
	\draw[fill=black] (axis cs:20,20)circle(0.075cm)node[right]{$N.E.$};
	\draw[fill=black] (axis cs:15,15)circle(0.075cm)node[above]{$Collusion$};
	\draw[fill=black] (axis cs:15,30)circle(0.075cm)node[right]{Stackelberg (S.A. leads)};
	\end{axis}
\end{tikzpicture}
\end{figure}

### Stackelberg Competition: Example V

\begin{example}

\end{example}

- With $q^*_{sa}=30$ and $q^*_i=15$, this sets a market price of
\vspace{-0.6cm}\begin{align*}
P&=200-3(45)\\
P&=65\\
\end{align*}\vspace{-0.6cm}
- Profit for Saudi Arabia is 
\vspace{-0.75cm}\begin{align*}
\pi_{sa}&=30(65-20)\\
\pi_{sa}&=\$1,350\\
\end{align*}\vspace{-0.6cm}
- Profit for Iran is 
\vspace{-0.6cm}\begin{align*}
\pi_{i}&=15(65-20)\\
\pi_{i}&=\$675\\
\end{align*}\vspace{-0.6cm}

### Simultaneous vs. Sequential Quantity Competition

\begin{table}
		\centering 
		\begin{tabular}{lrrrr}
	& \multicolumn{2}{c}{Cournot} & \multicolumn{2}{c}{Stackelberg}\\ 
	& \multicolumn{2}{c}{(P*=\$80)} & \multicolumn{2}{c}{(P*=\$65)}\\ \cmidrule{2-3} \cmidrule{4-5}
	Country & Output & Profit & Output & Profit\\ \toprule 
	Saudi Arabia & 20 & \$1,200 & 30 & \$1,350\\
	Iran & 20 & \$1,200 & 15 & \$675\\ \midrule 
	Industry & 40 & \$2,400 & 45 & \$2,025\\ \bottomrule 
	\end{tabular}
\end{table}

- 1^st^ mover Saudi Arabia increased its output and increased its profits
- 2^nd^ mover Iran was forced to lower its output and accept lower profits

### Stackelberg and First-Mover Advantage

\columnsbegin

\column{.5\textwidth}

- Stackelberg leader clearly has a **first-mover advantage** over the follower
    - Leader: $q^*=30$, $\pi=\$1,350$
    - Follower: $q^*=15$, $\pi=\$675$
    - If simultaneous (Cournot): $q^*=20$, $\pi=\$1,200$ each
    - Producing first $\succ$ producing simultaneously $\succ$ producing last

\column{.5\textwidth}

\begin{figure}
\centering
\begin{tikzpicture}\scriptsize 
	\begin{axis}[
		scale=0.8,
		axis lines=middle, 
		enlarge x limits={rel=0.1, upper},
		enlarge y limits={rel=0.1, upper},
		every axis y label/.style={at={(axis description cs:-0.2,0.5)},rotate=90,anchor=north},
		every axis x label/.style={at={(axis description cs:0.5,-0.1)},anchor=north},
	legend style={at={(0.5,-0.25)},anchor=north},
	xlabel=Iran's output ($q_i$),
	ylabel=Saudi Arabia's output ($q_{sa}$),
	xtick={0,10,...,60},
	ytick={0,10,...,60},
	grid=major,
	ymin=0,
	xmin=0,
	ymax=60,
	xmax=60,
	legend cell align=left]
	\addplot[domain=0:60, ultra thick, red]{30-0.5*x};
	\addlegendentry{Saudi Arabia's Reaction Curve}
	\addplot[domain=0:60, ultra thick, blue]{60-2*x};
	\addlegendentry{Iran's Reaction Curve}
	\draw[fill=black] (axis cs:20,20)circle(0.075cm)node[right]{$N.E.$};
	\draw[fill=black] (axis cs:15,15)circle(0.075cm)node[above]{$Collusion$};
	\draw[fill=black] (axis cs:15,30)circle(0.075cm)node[right]{Stackelberg (S.A. leads)};
	\end{axis}
\end{tikzpicture}
\end{figure}

\columnsend

### Stackelberg and First-Mover Advantage II 

\columnsbegin

\column{.5\textwidth}

- Stackelberg Nash equilibrium requires perfect information for *both* leader and follower
    - Follower must be able to *observe* leader's output to choose its own
    - Leader must *believe* follower will see leader's output and react optimally
- Imperfect information reduces the game to (simultaneous) Cournot competition

\column{.5\textwidth}

\begin{figure}
\centering
\begin{tikzpicture}\scriptsize 
	\begin{axis}[
		scale=0.8,
		axis lines=middle, 
		enlarge x limits={rel=0.1, upper},
		enlarge y limits={rel=0.1, upper},
		every axis y label/.style={at={(axis description cs:-0.2,0.5)},rotate=90,anchor=north},
		every axis x label/.style={at={(axis description cs:0.5,-0.1)},anchor=north},
	legend style={at={(0.5,-0.25)},anchor=north},
	xlabel=Iran's output ($q_i$),
	ylabel=Saudi Arabia's output ($q_{sa}$),
	xtick={0,10,...,60},
	ytick={0,10,...,60},
	grid=major,
	ymin=0,
	xmin=0,
	ymax=60,
	xmax=60,
	legend cell align=left]
	\addplot[domain=0:60, ultra thick, red]{30-0.5*x};
	\addlegendentry{Saudi Arabia's Reaction Curve}
	\addplot[domain=0:60, ultra thick, blue]{60-2*x};
	\addlegendentry{Iran's Reaction Curve}
	\draw[fill=black] (axis cs:20,20)circle(0.075cm)node[right]{$N.E.$};
	\draw[fill=black] (axis cs:15,15)circle(0.075cm)node[above]{$Collusion$};
	\draw[fill=black] (axis cs:15,30)circle(0.075cm)node[right]{Stackelberg (S.A. leads)};
	\end{axis}
\end{tikzpicture}
\end{figure}

\columnsend

### Stackelberg and First-Mover Advantage III 

\columnsbegin

\column{.5\textwidth}

- Leader cannot act like a monopolist 
    - A strategic game! Market output (that pushes down market price) is $Q=q_1+q_2$
- Leader's choice of 30 is optimal *only if* follower responds with 15

\column{.5\textwidth}

\begin{figure}
\centering
\begin{tikzpicture}\scriptsize 
	\begin{axis}[
		scale=0.8,
		axis lines=middle, 
		enlarge x limits={rel=0.1, upper},
		enlarge y limits={rel=0.1, upper},
		every axis y label/.style={at={(axis description cs:-0.2,0.5)},rotate=90,anchor=north},
		every axis x label/.style={at={(axis description cs:0.5,-0.1)},anchor=north},
	legend style={at={(0.5,-0.25)},anchor=north},
	xlabel=Iran's output ($q_i$),
	ylabel=Saudi Arabia's output ($q_{sa}$),
	xtick={0,10,...,60},
	ytick={0,10,...,60},
	grid=major,
	ymin=0,
	xmin=0,
	ymax=60,
	xmax=60,
	legend cell align=left]
	\addplot[domain=0:60, ultra thick, red]{30-0.5*x};
	\addlegendentry{Saudi Arabia's Reaction Curve}
	\addplot[domain=0:60, ultra thick, blue]{60-2*x};
	\addlegendentry{Iran's Reaction Curve}
	\draw[fill=black] (axis cs:20,20)circle(0.075cm)node[right]{$N.E.$};
	\draw[fill=black] (axis cs:15,15)circle(0.075cm)node[above]{$Collusion$};
	\draw[fill=black] (axis cs:15,30)circle(0.075cm)node[right]{Stackelberg (S.A. leads)};
	\end{axis}
\end{tikzpicture}
\end{figure}

\columnsend

### Comparing All Oligopoly Models

\begin{table}\scriptsize 
		\centering 
		\begin{tabular}{lrrrrrrrrrrrrrrr}
	& \multicolumn{3}{c}{Bertrand} & & \multicolumn{3}{c}{Cournot} & & \multicolumn{3}{c}{Stackelberg} & & \multicolumn{3}{c}{Collusion} \\ \cmidrule{2-4} \cmidrule{6-8} \cmidrule{10-12} \cmidrule{14-16}
	Country & $q$ & $p$ & $\pi$ & & $q$ & $p$ & $\pi$ & &$q$ & $p$ & $\pi$  & & $q$ & $p$ & $\pi$\\ \toprule 
	Saudi Arabia & 30 & \$20 & \$0 & & 20 & \$80 & \$1,200 & & 30 & \$65 & \$1,350 & & 15 & \$110 & \$1,400\\
	Iran & 30 & \$20 & \$0 & & 20 & \$80 & \$1,200 & & 15 & \$65 & \$675& & 15 & \$110 & \$1,400\\ \midrule 
	Industry & 60 & \$20 & \$0 & & 40 & \$80 & \$2,400 & & 45  & \$65 & \$2,025& & 30 & \$110 & \$2,800\\ \bottomrule 
	\end{tabular}
\end{table}

- Consumers do best under Bertrand competition $\succ$ Stackelberg $\succ$ Cournot $\succ$ Cartel
- Firm profits highest under cartel

### Comparing All Oligopoly Models II

\columnsbegin

\column{.5\textwidth}

- We can visualize \textcolor{blue}{consumer surplus}, \textcolor{teal}{profits}, and \textbf{deadweight loss} across the types of models

\column{.5\textwidth}

\begin{tikzpicture}\scriptsize 
	\begin{axis}[
		scale=0.8,
		axis lines=middle, 
		enlarge x limits={rel=0.1, upper},
		enlarge y limits={rel=0.1, upper},
		every axis y label/.style={at={(axis description cs:-0.2,0.5)},rotate=90,anchor=north},
		every axis x label/.style={at={(axis description cs:0.5,-0.1)},anchor=north},
	xlabel=Industry Output (Q),
	ylabel=Market Price (p),
	xtick={0, 10,20,...,100},
	ytick={0,20,40,...,200},
	legend style={at={(0.5,-0.25)},anchor=north},
	grid=major,
	ymin=0,
	xmin=0,
	ymax=200,
	xmax=100,
	legend cell align=left,
]
	\addplot[domain=0:100,ultra thick, red!50]{20};
	\addlegendentry{$MC(q)=AC(q)$}
	\addplot[domain=0:100,ultra thick,blue!50]{200-3*x};
	\addlegendentry{Demand}
	\addplot[domain=0:100,ultra thick,purple!50]{200-6*x};
	\addlegendentry{$MR(q)$}
	\draw<2->[fill=black] (axis cs:60,20)circle(0.075cm)node[right]{$Bertrand (PC)$};
	\draw<2->[dashed, thick, black] (axis cs:0,20)--(axis cs:60,20)--(axis cs:60,0);
	\draw<3->[fill=black] (axis cs:40,80)circle(0.075cm)node[right]{$Cournot$};
	\draw<3->[dashed, thick, black] (axis cs:0,80)--(axis cs:40,80)--(axis cs:40,0);
	\draw<4->[fill=black] (axis cs:45,65)circle(0.075cm)node[right]{$Stackelberg$};
	\draw<4->[dashed, thick, black] (axis cs:0,65)--(axis cs:45,65)--(axis cs:45,0);
	\draw<5->[fill=black] (axis cs:30,110)circle(0.075cm)node[right]{$Cartel$};
	\draw<5->[dashed, thick, black] (axis cs:0,110)--(axis cs:30,110)--(axis cs:30,0);
\only<2>{\addplot[fill=blue, fill opacity = 0.2, draw = none,area legend] coordinates {(0,20) (0,200) (60,20)};
\addlegendentry{Consumer Surplus};}
\only<3>{\addplot[fill=blue, fill opacity = 0.2, draw = none,area legend] coordinates {(0,80) (0,200) (40,80)};
\addlegendentry{Consumer Surplus};}
\only<3>{\addplot[fill=teal, fill opacity = 0.5, draw = none,area legend] coordinates {(0,20) (0,80) (40,80) (40,20)};
\addlegendentry{Industry Profit};}
\only<3>{\addplot[fill=black, fill opacity = 0.5, draw = none,area legend] coordinates {(40,20) (40,80) (60,20)};
\addlegendentry{Deadweight Loss};}
\only<4>{\addplot[fill=blue, fill opacity = 0.2, draw = none,area legend] coordinates {(0,65) (0,200) (45,65)};
\addlegendentry{Consumer Surplus};}
\only<4>{\addplot[fill=teal, fill opacity = 0.5, draw = none,area legend] coordinates {(0,20) (0,65) (45,65) (45,20)};
\addlegendentry{Industry Profit};}
\only<4>{\addplot[fill=black, fill opacity = 0.5, draw = none,area legend] coordinates {(45,20) (45,65) (60,20)};
\addlegendentry{Deadweight Loss};}
\only<5>{\addplot[fill=blue, fill opacity = 0.2, draw = none,area legend] coordinates {(0,110) (0,200) (30,110)};
\addlegendentry{Consumer Surplus};}
\only<5>{\addplot[fill=teal, fill opacity = 0.5, draw = none,area legend] coordinates {(0,20) (0,110) (30,110) (30,20)};
\addlegendentry{Industry Profit};}
\only<5>{\addplot[fill=black, fill opacity = 0.5, draw = none,area legend] coordinates {(30,20) (30,110) (60,20)};
\addlegendentry{Deadweight Loss};}

        \end{axis}
\end{tikzpicture}	

\columnsend

### Stackelberg Competition, You Try

\begin{example}
Continuing from before, Firm 1 and Firm 2 have a constant $MC=AC=\$8$. The market (inverse) demand curve is given by:
\vspace{-0.5cm}\begin{align*}
P&=200-2Q\\
Q&=q_1+q_2\\
\end{align*}\vspace{-0.5cm}

d. Suppose now Firm 1 is a Stackelberg leader. Find each firm's output and profit. 

\end{example}

### Stackelberg Competition Game (Moblab)

\begin{figure}
  \includegraphics[height=2.75in]{stackelbergmoblab}
\end{figure}

### Stackelberg Competition Game (Moblab) II 

\columnsbegin

\column{.7\textwidth}

- Each of you is one Airline competing against another in a duopoly
    - Each pays same per-flight cost
    - Market price determined by *total* number of flights in market
- **LeadAir** first chooses its number of flights, publicly announced
- **FollowAir** then chooses its number of flights

\column{.3\textwidth}

\begin{figure}
  \includegraphics[height=1.25in]{stackelbergmoblab}
\end{figure}

\columnsend

## Contestable Markets

### Is Monopoly a Nash Equilibrium?

\columnsbegin

\column{.7\textwidth}

- Are outcomes of other market structures Nash equilibria?
- **Perfect competition**: no firm wants to raise or lower price given market price
- **Monopolist** maximizes $\pi$ by setting $q^*$: $MR=MC$ and $p^*=Demand(q^*)$
- This is *an* equilibrium, but is it the *only* equilibrium?
    - What if customers refuse to pay $p>MC$?
- Ignores (or assumes away) *potential* competition

\column{.3\textwidth}

\begin{figure}
  \includegraphics[width=1.5in]{profitloss}
  \includegraphics[width=1.5in]{enterexit}
\end{figure}

\columnsend

### Contestable Markets

\columnsbegin

\column{.7\textwidth}

- Model the market as an entry game, with an **Incumbent** who sets $p_i$
- **Entrant** decides to **stay out** or **enter** the market, charging $p_e$
- Consumers buy from the firm with the lower price
- Bertrand-style price competition

\column{.3\textwidth}

\begin{figure}
  \includegraphics[width=1.5in]{profitloss}
  \includegraphics[width=1.5in]{enterexit}
\end{figure}

\columnsend

### Contestable Markets II

\columnsbegin

\column{.45\textwidth}

- Suppose firms have a total cost of $C(q)=cq$
    - Thus, a marginal cost of $MC(q)=c$
- If Incumbent sets $p_i>c$, then entrant would enter and set $p_e=p_i-\epsilon$
- **Nash Equilibrium**: incumbent sets $p_i=c$, no entry
    - "Monopoly": a single firm, but competitive outcomes!
    - $p=MC$, $\pi=0$, max Consumer Surplus

\column{.55\textwidth}

```{r} 
#, fig.height=4.5, out.width='\\textwidth', fig.asp=0.75
### Input parameters

# Demand function, format: p=d-eq
a<-10
b<-2
f<-5
c<-2

Demand=function(x){a-b*x}
MR=function(x){a-2*b*x}
MC=function(x){c}
AC=function(x){f/x+c}

q_choke=a/b
q_PC<-(a-c)/(b)

CS<-data.frame(x=c(0,0,q_PC),
               y=c(Demand(q_PC),a,Demand(q_PC)))

line_colors <- c("Demand" = "blue", "MR" = "purple", "MC" = "red", "AC" = "orange")
surplus_colors <-c("Consumer" = "blue", "Profit" = "green", "DWL" = "black")

# make plot 

library("ggplot2")
ggplot(data.frame(x=c(0,q_choke)), aes(x=x))+
  
  # curves 
  stat_function(fun=Demand, geom="line", size=1, aes(color="Demand"))+
  stat_function(fun=MR, geom="line", size=1, aes(color="MR"))+
  stat_function(fun=MC, geom="line", size=1, aes(color="MC"))+
  # surpluses
  geom_polygon(data=CS, aes(x=x,y=y, fill="Consumer"), alpha=0.5)+ #CS

  # labels 
  geom_segment(aes(x=q_PC,y=0),xend=q_PC,yend=c,linetype=3)+
  geom_point(aes(x=q_PC,y=c), color="black", size=3)+ 
  geom_text(aes(x=q_PC,y=c,label="N.E.", family="Yanone Kaffeesatz Regular"),size = 9,hjust=-0.1, vjust=-0.25, color="black")+
  annotate(geom = "label", x = 4.5, y = Demand(4.5), label = "Demand",
           color = "blue", size = 9) +
  annotate(geom = "label", x = 2.25, y = MR(2.25), label = "MR",
           color = "purple", size = 9) +
  annotate(geom = "label", x = 4.5, y = c, label = "MC",
           color = "red", size = 9) +
  annotate(geom = "label", x = 1.3, y = 5, label = "Consumer Surplus",
           color = "blue", size = 9) +
  scale_x_continuous(breaks = c(0,q_PC),
                     labels = c(0,expression(q[p=MC])),
                     expand=c(0,0))+
  scale_y_continuous(breaks = c(0,c,a),
                     labels = c(0,expression(c),expression(a)),
                     limits = c(0,a),
                     expand=c(0,0))+
  theme_light()+
  scale_colour_manual("Curves", values = line_colors)+
  scale_fill_manual("Surpluses", values = surplus_colors, guide = guide_legend(override.aes = list(alpha = 0.5)))+
  labs(x = "Quantity (q)",
       y = "Price (p)")+
  theme(text=element_text(size=30, family="Yanone Kaffeesatz Regular"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "")
```

\columnsend

### Contestable Markets III

\columnsbegin

\column{.4\textwidth}

- What if entrant has higher costs: $c_e>c_i$? 
    - **Nash equilibrium**: incumbent sets $p_i=p_e-\epsilon$
    - No entry
- A monopoly, but not worst case

\column{.6\textwidth}

```{r} 
#fig.asp=0.75
### Input parameters

# Demand function, format: p=d-eq
a<-10
b<-2
f<-5
c<-2
d<-4.5
e<-4

Demand=function(x){a-b*x}
MR=function(x){a-2*b*x}
MCi=function(x){c}
MCe=function(x){d}
AC=function(x){f/x+c}

q_choke=a/b
q_PC<-(a-c)/(b)

q_star<-(a-e)/b
q_monopoly<-(a-c)/(2*b)

CS<-data.frame(x=c(0,0,q_star),
               y=c(Demand(q_star),a,Demand(q_star)))
DWL<-data.frame(x=c(q_star,q_star,q_PC),
               y=c(c,e,c))

line_colors <- c("Demand" = "blue", "MR" = "purple", "MCi" = "red", "MCe" = "red4")
surplus_colors <-c("Consumer" = "blue", "Profit" = "green", "DWL" = "black")

# make plot 

library("ggplot2")
ggplot(data.frame(x=c(0,q_choke)), aes(x=x))+
  
  # curves 
  stat_function(fun=Demand, geom="line", size=1, aes(color="Demand"))+
  stat_function(fun=MR, geom="line", size=1, aes(color="MR"))+
  stat_function(fun=MCi, geom="line", size=1, aes(color="MCi"))+
  stat_function(fun=MCe, geom="line", size=1, aes(color="MCe"))+
  # surpluses
  geom_polygon(data=CS, aes(x=x,y=y, fill="Consumer"), alpha=0.5)+ #CS
  geom_rect(aes(xmin=0,xmax=q_star,ymin=c,ymax=e,fill="Profit"),alpha=0.25) + # Profits
  geom_polygon(data=DWL, aes(x=x,y=y, fill="DWL"), alpha=0.5)+ #CS
  # labels 
  
  # P.C. point 
  geom_segment(aes(x=q_PC,y=0),xend=q_PC,yend=c,linetype=3)+
  geom_point(aes(x=q_PC,y=c), color="black", size=3)+ 
  geom_text(aes(x=q_PC,y=c,label="P.C.", family="Yanone Kaffeesatz Regular"),size = 9,hjust=-0.1, vjust=-0.25, color="black")+
  
  # N.E. point 
  geom_segment(aes(x=q_star,y=0),xend=q_star,yend=e,linetype=3)+
  geom_segment(aes(x=0,y=e),xend=q_star,yend=e,linetype=3)+
  geom_point(aes(x=q_star,y=e), color="black", size=3)+ 
  geom_text(aes(x=q_star,y=e,label="N.E.", family="Yanone Kaffeesatz Regular"),size = 9,hjust=-0.1, vjust=-0.25, color="black")+
  
  # Monopoly point 
  geom_segment(aes(x=q_monopoly,y=0),xend=q_monopoly,yend=Demand(q_monopoly),linetype=3)+
  geom_segment(aes(x=0,y=Demand(q_monopoly)),xend=q_monopoly,yend=Demand(q_monopoly),linetype=3)+
  geom_point(aes(x=q_monopoly,y=Demand(q_monopoly)), color="black", size=3)+ 
  geom_text(aes(x=q_monopoly,y=Demand(q_monopoly),label="Monopoly", family="Yanone Kaffeesatz Regular"),size = 9, hjust=-0.1, vjust=-0.25, color="black")+
  annotate(geom = "label", x = 4.5, y = Demand(4.5), label = "Demand",
           color = "blue", size = 9) +
  annotate(geom = "label", x = 2.25, y = MR(2.25), label = "MR",
           color = "purple", size = 9) +
  annotate(geom = "label", x = 4.5, y = c, label = expression(MC[i]),
           color = "red", size = 9) +
  annotate(geom = "label", x = 4.5, y = d, label = expression(MC[e]),
           color = "red4", size = 9) +
  annotate(geom = "label", x = 1.25, y = 6, label = "Consumer Surplus",
           color = "blue", size = 9) +
  annotate(geom = "label", x = 1.25, y = 3, label = "Profit",
           color = "green", size = 9) +
  annotate(geom = "label", x = 3.2, y = 3, label = "DWL",
           color = "black", size = 9) +
  scale_x_continuous(breaks = c(0,q_monopoly,q_star,q_PC),
                     labels = c(0,expression(q[monopoly]),expression(q[NE]),expression(q[p=MC])),
                     expand=c(0,0))+
  scale_y_continuous(breaks = c(0,c,d,e,Demand(q_monopoly),a),
                     labels = c(0,expression(MC[i]),expression(MC[e]),expression(p[i]==MC[e]-epsilon),expression(p[monopoly]),expression(a)),
                     limits = c(0,a),
                     expand=c(0,0))+
  theme_light()+
  scale_colour_manual("Curves", values = line_colors)+
  scale_fill_manual("Surpluses", values = surplus_colors, guide = guide_legend(override.aes = list(alpha = 0.5)))+
  labs(x = "Quantity (q)",
       y = "Price (p)")+
  theme(text=element_text(size=30, family="Yanone Kaffeesatz Regular"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "")
```

\columnsend

### Contestable Markets IV

\columnsbegin

\column{.5\textwidth}

- Suppose firms have the same cost structure again
- What if there are *fixed costs*, $f$?
    - $C(q)=f+cq$
    - $MC(q)=c$
    - $AC(q)=c+\frac{f}{q}$
- Economies of scale prevent marginal cost pricing from being profitable
    - $\pi=-\frac{f}{q}<0$

\column{.5\textwidth}

```{r, fig.height=4, fig.asp=0.75}
### Input parameters

library("mosaic")

# Demand function, format: p=d-eq
a<-10
b<-2
f<-5
c<-2

Demand=function(x){a-b*x}
MR=function(x){a-2*b*x}
MC=function(x){c}
AC=function(x){f/x+c}

q_choke<-(a/b)
MR_choke<-(a/(2*b))
q_PC<-(a-c)/(b)

CS<-data.frame(x=c(0,0,q_PC),
               y=c(Demand(q_PC),a,Demand(q_PC)))

line_colors <- c("Demand" = "blue", "MR" = "purple", "MC" = "red", "AC" = "orange")
surplus_colors <-c("Consumer" = "blue", "Loss" = "black", "DWL" = "black")

# make plot 

library("ggplot2")
ggplot(data.frame(x=c(0,q_choke)), aes(x=x))+
  
  # curves 
  stat_function(fun=Demand, geom="line", size=1, aes(color="Demand"))+
  stat_function(fun=MR, geom="line", size=1, aes(color="MR"))+
  stat_function(fun=MC, geom="line", size=1, aes(color="MC"))+
  stat_function(fun=AC, geom="line", size=1, aes(color="AC"))+
  
  # surpluses
  geom_rect(aes(xmin=0,xmax=q_PC,ymin=c,ymax=AC(q_PC),fill="Loss"),alpha=0.25) + # Losses
  
   # labels 
  geom_segment(aes(x=q_PC,y=0),xend=q_PC,yend=c,linetype=3)+
  geom_point(aes(x=q_PC,y=c), color="black", size=3)+ 
  geom_text(aes(x=q_PC,y=c,label="P.C.",family = "Yanone Kaffeesatz Regular"),size = 9,hjust=-0.1,
            vjust=-0.1, color="black")+
  annotate(geom = "label", x = 4.5, y = Demand(4.5), label = "Demand",
           color = "blue", size = 9) +
  annotate(geom = "label", x = 2.25, y = MR(2.25), label = "MR",
           color = "purple", size = 9) +
  annotate(geom = "label", x = 4.5, y = c, label = "MC",
           color = "red", size = 9) +
  annotate(geom = "label", x = 4.5, y = AC(d), label = "AC",
           color = "orange", size = 9) +
  annotate(geom = "label", x = 1.25, y = 2.5, label = "Loss",
           color = "black", size = 9) +
  scale_x_continuous(breaks = c(0,q_PC),
                     labels = c(0,expression(q[p=MC])),
                     expand=c(0,0))+
  scale_y_continuous(breaks = c(0,c,AC(q_PC),a),
                     labels = c(0,expression(c),expression(c+f/q),expression(a)),
                     limits = c(0,a),
                     expand=c(0,0))+
  theme_light(base_family = "Yanone Kaffeesatz Regular")+
  scale_colour_manual("Curves", values = line_colors)+
  scale_fill_manual("Surpluses", values = surplus_colors, guide = guide_legend(override.aes = list(alpha = 0.5)))+
  labs(x = "Quantity (q)",
       y = "Price (p)")+
    theme(text=element_text(size=30, family="Yanone Kaffeesatz Regular"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          legend.position = "")
```

\columnsend

### Contestable Markets IV

\columnsbegin

\column{.5\textwidth}

- *Nash equilibrium*: Incumbent prices at $p_i=AC$ earns $\pi=0$
- No entry
- Again, "monopoly" but earning no profits

\column{.5\textwidth}

```{r, fig.height=4, fig.asp=0.75}
### Input parameters

library("mosaic")

# Demand function, format: p=d-eq
a<-10
b<-2
f<-5
c<-2

Demand=function(x){a-b*x}
MR=function(x){a-2*b*x}
MC=function(x){c}
AC=function(x){f/x+c}

q_choke<-(a/b)
MR_choke<-(a/(2*b))
q_PC<-(a-c)/(b)
q_AC<-((a-c)+sqrt((a-c)^2-4*b*f))/(2*b) # needed quadratic formula, heh 

CS<-data.frame(x=c(0,0,q_AC),
               y=c(Demand(q_AC),a,Demand(q_AC)))

DWL<-data.frame(x=c(q_AC,q_AC,q_PC),
                        y=c(c,Demand(q_AC),c))

line_colors <- c("Demand" = "blue", "MR" = "purple", "MC" = "red", "AC" = "orange")
surplus_colors <-c("Consumer" = "blue", "Profit" = "green", "DWL" = "black")

# make plot 

library("ggplot2")
ggplot(data.frame(x=c(0,q_choke)), aes(x=x))+
  
  # curves 
  stat_function(fun=Demand, geom="line", size=1, aes(color="Demand"))+
  stat_function(fun=MR, geom="line", size=1, aes(color="MR"))+
  stat_function(fun=MC, geom="line", size=1, aes(color="MC"))+
  stat_function(fun=AC, geom="line", size=1, aes(color="AC"))+
  
  # surpluses
  geom_polygon(data=CS, aes(x=x,y=y, fill="Consumer"), alpha=0.5)+ #CS
  geom_polygon(data=DWL, aes(x=x,y=y, fill="DWL"), alpha=0.5)+ #DWL
  
   # labels 
  geom_segment(aes(x=q_PC,y=0),xend=q_PC,yend=c,linetype=3)+
  geom_point(aes(x=q_PC,y=c), color="black",size=3)+ 
  geom_text(aes(x=q_PC,y=c,label="P.C.",family = "Yanone Kaffeesatz Regular"),size = 9,hjust=-0.1,
            vjust=-0.1, color="black")+
  geom_segment(aes(x=q_AC,y=0),xend=q_AC,yend=Demand(q_AC), linetype=3)+
  geom_segment(aes(x=0,y=Demand(q_AC)),xend=q_AC,yend=Demand(q_AC), linetype=3)+
  geom_text(aes(x=q_AC,y=Demand(q_AC),label="N.E.",family = "Yanone Kaffeesatz Regular"),size = 9,hjust=-0.1,
            vjust=-0.1, color="black")+
  annotate(geom = "label", x = 4.5, y = Demand(4.5), size = 9, label = "Demand",
           color = "blue") +
  annotate(geom = "label", x = 2.25, y = MR(2.25), size = 9, label = "MR",
           color = "purple") +
  annotate(geom = "label", x = 4.5, y = c, size = 9, label = "MC",
           color = "red") +
  annotate(geom = "label", x = 4.5, y = AC(d), size = 9, label = "AC",
           color = "orange") +
  annotate(geom = "label", x = 1.25, y = 6, label = "Consumer Surplus",
           color = "blue", size = 9) +
  annotate(geom = "label", x = 3.5, y = 2.5, label = "DWL",
           color = "black", size = 9) +
  scale_x_continuous(breaks = c(0,q_AC,q_PC),
                     labels = c(0,expression(q[p=AC]),expression(q[p=MC])),
                     expand=c(0,0))+
  scale_y_continuous(breaks = c(0,c,AC(q_AC),a),
                     labels = c(0,expression(c),expression(c+f/q),expression(a)),
                     limits = c(0,a),
                     expand=c(0,0))+
  theme_light(base_size=22,base_family = "Yanone Kaffeesatz Regular")+
  scale_colour_manual("Curves", values = line_colors)+
  scale_fill_manual("Surpluses", values = surplus_colors, guide = guide_legend(override.aes = list(alpha = 0.5)))+
  labs(x = "Quantity (q)",
       y = "Price (p)")+
    theme(text=element_text(size=30, family="Yanone Kaffeesatz Regular"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          legend.position = "")
```

\columnsend

### What About *Sunk* Costs?

\columnsbegin

\column{.5\textwidth}

- Fixed costs $\implies$ do not vary with output
- If firm exits, could sell these assets (e.g. machines, real estate) to recover costs
    - Thus, "hit-and-run" competition remains potentially profitable
    - Maintains credible threat against incumbent acting as a monopolist
- But what if assets are *not* sellable and costs *not* recoverable---i.e. **sunk costs**?
    - e.g. research and development, spending to build brand equity, advertising, worker-training for industry-specific skills, etc 

\column{.5\textwidth}

\begin{figure}
  \includegraphics[height=1.5in]{sunkcost}
\end{figure}

\columnsend

### What About *Sunk* Costs? II 

\columnsbegin

\column{.5\textwidth}

- These are bygones to the incumbent, who has already committed to producing
- But new costs and risk to entrant, lowering expected profits
- In effect, sunk costs make $c_e>e_i$
- Incumbent can deter entry with $p_i=p_e-\epsilon$
- Inefficient $p>AC$, but again lower than monopoly level

\column{.5\textwidth}

\begin{figure}
  \includegraphics[height=1.5in]{sunkcost}
\end{figure}

\columnsend

### Contestable Markets: Recap

- Markets are contestable if:
    - There are no barriers to entry or exit
    - Firms have similar technologies (i.e. similar cost structure)
    - There are no sunk costs
- Economies of scale need not be inconsistent with competitive markets (as is assumed) if they are contestable
- Generalizes "prefect competition" model in more realistic way, also game-theoretic

### Contestable Markets VI

\columnsbegin

\column{.7\textwidth}

> "This means that...an incumbent, even if he can threaten retaliation after entry, dare not offer profit-making opportunities to potential entrants because an entering firm can hit and run, gathering in the available profits and departing when the going gets rough." 

Baumol, William, J, 1982, "Contestable Markets: An Uprising in the Theory of Industry Structure," *American Economic Review,* 72(1): 1-15

\column{.3\textwidth}

\begin{tabular}{c}
\includegraphics[height=2in]{baumol1}\\
William J. Baumol\\
(1922--2017)\\
\end{tabular}

\columnsend

### Implications 

- "Old" industrial organization (pre-1980s)
    - Structure$\implies$Conduct$\implies$Performance (SCP) paradigm
- Number of firms in an industry is **exogenous**
    - Determined solely by unchanging parameters: technology, costs, demand
- Regulation & antitrust focus on number of firms
    - "Count the number of firms, if it's 1, it's a monopoly!"
- Perfect competition as gold standard, only market arrangement that is socially efficient
    - $p=MC$, $\pi_{LR}=0$

### Implications II

- The number of firms in an industry is **endogenous**
    - Not assumed exogenously, but a function of how firms respond game-theoretically to each other
- Perfect competition not the only market-structure that is socially efficient
    - Monopoly and oligopoly can be efficient if they are contestable
- Regulation and antitrust should consider whether a market is *contestable*, not just the *number* of firms
    - Free entry
    - No sunk costs

### Implications III

- Industries with one firm *may* not be acting like monopolies!
    - Set $p=c$ or $p=AC$ with $\pi=0$ to prevent entry
- Entry will still bid prices to competitive levels for consumers, erode profits for firms
- Firms engaging in egregious monopoly behavior ($\downarrow q$, $\uparrow p$, $\pi>0$) only persist because of **barriers to entry**
    - Activities or institutions to raise $b_e>b_i$


### Implications IV

\columnsbegin

\column{.6\textwidth}

> "Of far greater concern to Microsoft is the competition from new and emerging technologies, some of which are currently visible and others of which certainly are not. This array of known, emerging, and wholly unknown competitors places enormous pressure on Microsoft to price competitively and innovate aggressively." (Schmalensee 1999)

United States v. Microsoft (1998)

\column{.4\textwidth}

\begin{figure}
  \includegraphics[height=2in]{microsoftdeposition}
\end{figure}

\columnsend

# Broadening Solution Concepts

### Multiple Equilibria: What to Do?

\columnsbegin

\column{.6\textwidth}

- Nash equilibrium is the most well known **solution concept** in game theory
    - Method of predicting the outcome of a game
- Suppose we have a coordination game with multiple Nash equilibria
- What can we say about the behavior of players?
- One answer: nothing!
    - Both equilibria are mutual best responses
    - Coordination problem on which strategy to jointly select
    - Two sides of the road to drive on, one no better than other

\column{.4\textwidth}

\begin{figure}
  \includegraphics[height=1in]{equilibriumbalance}
\end{figure}

\columnsend

### Multiple Equilibria: What to Do? II

\columnsbegin

\column{.6\textwidth}

- Another answer: we must confront multiple equilibria in economics
    - still want to predit *which* outcome will happen
- We need to consider multiple criteria beyond best responses to select an equilibrium
    - Focalness/salience
    - Fairness/envy-free-ness
    - Efficiency/payoff dominance
    - Risk dominance

\column{.4\textwidth}

\begin{figure}
  \includegraphics[height=1in]{equilibriumbalance}
\end{figure}

\columnsend

### Multiple Equilibria: Efficiency

\columnsbegin

\column{.6\textwidth}

- Which equilibrium is most **(Pareto) efficient?**
    - Must be no other equilibrium where *at least one* player earns a higher payoff and no player earns a lower payoff
- Stag Hunt game:
    - Both (\textcolor{red}{Stag}, \textcolor{blue}{Stag}) and (\textcolor{red}{Hare}, \textcolor{blue}{Hare}) are Nash equilibria
    - (\textcolor{red}{Stag}, \textcolor{blue}{Stag}) is Pareto superior to (\textcolor{red}{Hare}, \textcolor{blue}{Hare})

\column{.4\textwidth}

\begin{figure}
\centering 
\begin{tabular}{lrcc|cc|}
& & \multicolumn{4}{c} {\textbf{\textcolor{blue}{Player 2}}} \\
& & \multicolumn{2}{c}{\textcolor{blue}{Stag}} & \multicolumn{2}{c}{\textcolor{blue}{Hare}} \\ \cline{3-6}
\parbox[t]{2mm}{\multirow{4}{*}{\rotatebox[origin=c]{90}{\textbf{\textcolor{red}{Player 1}}}}}  & 
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Stag}}} & \textcolor{red}{$2$} & & \textcolor{red}{$0$} & \\ 
&  &\multicolumn{1}{|c}{} & \textcolor{blue}{$2$} & & \textcolor{blue}{$1$}\\ \cline{3-6}
&
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Hare}}} & \textcolor{red}{$1$} &  & \textcolor{red}{$1$} & \\ 
& & \multicolumn{1}{|c}{} & \textcolor{blue}{$0$} &  & \textcolor{blue}{$1$} \\ \cline{3-6}
\end{tabular}
\end{figure}

\columnsend

### Multiple Equilibria: Efficiency II

\columnsbegin

\column{.6\textwidth}

- Consider the "Pittsburgh left" game
- Two Nash equilibria: (\textcolor{red}{Left}, \textcolor{blue}{Yield}) and (\textcolor{red}{Yield}, \textcolor{blue}{Straight})
    - Each driver *prefers* that the other yield
- This is just a variant of **Chicken**
- Both equilibria are Pareto efficient!

\begin{center}
\includegraphics[height=1.75in]{pittsburghleft.jpeg}
\end{center}

\column{.4\textwidth}

\begin{figure}
\centering 
\begin{tabular}{lrcc|cc|}
& & \multicolumn{4}{c} {\textbf{\textcolor{blue}{Driver 2}}} \\
& & \multicolumn{2}{c}{\textcolor{blue}{Straight}} & \multicolumn{2}{c}{\textcolor{blue}{Yield}} \\ \cline{3-6}
\parbox[t]{2mm}{\multirow{4}{*}{\rotatebox[origin=c]{90}{\textbf{\textcolor{red}{Row}}}}}  & 
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Left}}} & \textcolor{red}{$-100$} & & \textcolor{red}{\underline{5}} & \\ 
&  &\multicolumn{1}{|c}{} & \textcolor{blue}{$-100$} & & \textcolor{blue}{\underline{$-5$}} \\ \cline{3-6}
&
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Yield}}} & \textcolor{red}{\underline{$-5$}} &  & \textcolor{red}{$-10$} & \\ 
& & \multicolumn{1}{|c}{} & \textcolor{blue}{\underline{5}} &  & \textcolor{blue}{$-10$} \\ \cline{3-6}
\end{tabular}
\end{figure} 

\columnsend

### Multiple Equilibria: Efficiency III

\columnsbegin

\column{.6\textwidth}

- Often have multiple equilibria
- Sometimes governments step in to enforce a single equilibrium 

\onslide<2->\begin{center}
\includegraphics[height=1.75in]{leftyield}
\end{center}

\column{.4\textwidth}

\begin{figure}
\centering 
\begin{tabular}{lrcc|cc|}
& & \multicolumn{4}{c} {\textbf{\textcolor{blue}{Driver 2}}} \\
& & \multicolumn{2}{c}{\textcolor{blue}{Straight}} & \multicolumn{2}{c}{\textcolor{blue}{Yield}} \\ \cline{3-6}
\parbox[t]{2mm}{\multirow{4}{*}{\rotatebox[origin=c]{90}{\textbf{\textcolor{red}{Row}}}}}  & 
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Left}}} & \textcolor{red}{$-100$} & & \textcolor{red}{\underline{5}} & \\ 
&  &\multicolumn{1}{|c}{} & \textcolor{blue}{$-100$} & & \textcolor{blue}{\underline{$-5$}} \\ \cline{3-6}
&
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Yield}}} & \textcolor{red}{\underline{$-5$}} &  & \textcolor{red}{$-10$} & \\ 
& & \multicolumn{1}{|c}{} & \textcolor{blue}{\underline{5}} &  & \textcolor{blue}{$-10$} \\ \cline{3-6}
\end{tabular}
\end{figure} 

\columnsend

### Multiple Equilibria: Risk Dominance

\columnsbegin

\column{.6\textwidth}

- Back to Stag Hunt with 2 PSNE: 
- (\textcolor{red}{Stag}, \textcolor{blue}{Stag}) is **efficient** and "**payoff dominant**"
    - Highest payoff for each player, no possible Pareto improvements
- (\textcolor{red}{Hare}, \textcolor{blue}{Hare}) is **"risk dominant"**
    - *Less riskier* equilibrium
    - By playing $Hare$, each player guarantees themself 1 regardless of the other player's strategy
    
\column{.4\textwidth}

\begin{figure}
\centering 
\begin{tabular}{lrcc|cc|}
& & \multicolumn{4}{c} {\textbf{\textcolor{blue}{Player 2}}} \\
& & \multicolumn{2}{c}{\textcolor{blue}{Stag}} & \multicolumn{2}{c}{\textcolor{blue}{Hare}} \\ \cline{3-6}
\parbox[t]{2mm}{\multirow{4}{*}{\rotatebox[origin=c]{90}{\textbf{\textcolor{red}{Player 1}}}}}  & 
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Stag}}} & \textcolor{red}{$2$} & & \textcolor{red}{$0$} & \\ 
&  &\multicolumn{1}{|c}{} & \textcolor{blue}{$2$} & & \textcolor{blue}{$1$}\\ \cline{3-6}
&
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Hare}}} & \textcolor{red}{$1$} &  & \textcolor{red}{$1$} & \\ 
& & \multicolumn{1}{|c}{} & \textcolor{blue}{$0$} &  & \textcolor{blue}{$1$} \\ \cline{3-6}
\end{tabular}
\end{figure}

\columnsend

## Beliefs and Rationalizability

### The Role of Beliefs, Revisited

\columnsbegin

\column{.6\textwidth}

- What about the following game?  
- \textcolor{blue}{Column} has a dominant strategy to always play \textcolor{blue}{Left}
- Given this, \textcolor{red}{Row} should play \textcolor{red}{Down}
- Unique Nash equilibrium: (\textcolor{red}{Down}, \textcolor{blue}{Left})
- If you were \textcolor{red}{Row}, would you risk playing \textcolor{red}{Down} if there was the slightest chance you think \textcolor{blue}{Column} would play \textcolor{blue}{Right}?

\column{.4\textwidth}

\begin{figure}
\centering 
\begin{tabular}{lrcc|cc|}
& & \multicolumn{4}{c} {\textbf{\textcolor{blue}{Column}}} \\
& & \multicolumn{2}{c}{\textcolor{blue}{Left}} & \multicolumn{2}{c}{\textcolor{blue}{Right}} \\ \cline{3-6}
\parbox[t]{2mm}{\multirow{4}{*}{\rotatebox[origin=c]{90}{\textbf{\textcolor{red}{Row}}}}}  & 
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Up}}} & \textcolor{red}{9} & & \textcolor{red}{\underline{8}} & \\ 
&  &\multicolumn{1}{|c}{} & \textcolor{blue}{\underline{10}} & & \textcolor{blue}{9.9} \\ \cline{3-6}
&
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Down}}} & \textcolor{red}{\underline{10}} &  & \textcolor{red}{$-1000$} & \\ 
& & \multicolumn{1}{|c}{} & \textcolor{blue}{\underline{10}} &  & \textcolor{blue}{9.9} \\ \cline{3-6}
\end{tabular}
\end{figure} 

\columnsend

### Rationalizability

- Nash equilibrium requires players to have accurate beliefs about the others' actions
- In equilibrium:
    1. Each player should choose their highest-payoff strategy given their beliefs about other players' strategies
    2. The beliefs should be correct, i.e. match what other players actually do
- **Rationalizable** outcomes are a more general solution concept than Nash equilibria, allow for variations in beliefs

### Rationalizability

\columnsbegin

\column{.6\textwidth}

- Consider the following game, using best response analysis
- Unique Nash equilibrium: (\textcolor{red}{Middle}, \textcolor{blue}{Middle})
- \textcolor{red}{Row} plays \textcolor{red}{Middle} because she believes \textcolor{blue}{Column} will rationally play \textcolor{blue}{Middle} (who plays that because he believes \textcolor{red}{Row} will rationally play \textcolor{red}{Middle})

\column{.4\textwidth}

\begin{figure}
\centering
\begin{tabular}{llcc|cc|cc|}
& & \multicolumn{6}{c} {\textbf{\textcolor{blue}{Column}}} \\
& & \multicolumn{2}{c}{\textcolor{blue}{Left}} & \multicolumn{2}{c}{\textcolor{blue}{Middle}} & \multicolumn{2}{c}{\textcolor{blue}{Right}} \\ \cline{3-8}
\parbox[t]{2mm}{\multirow{6}{*}{\rotatebox[origin=c]{90}{\textbf{\textcolor{red}{Row}}}}}  & 
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Left}}} & \textcolor{red}{$0$} & & \textcolor{red}{$2$} & & \textcolor{red}{$\underline{7}$} & \\ 
&  &\multicolumn{1}{|c}{} & \textcolor{blue}{$\underline{7}$} & & \textcolor{blue}{$5$} & & \textcolor{blue}{$0$}\\ \cline{3-8}
&
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Middle}}} & \textcolor{red}{$5$} &  & \textcolor{red}{$\underline{3}$} & & \textcolor{red}{$5$} & \\ 
& & \multicolumn{1}{|c}{} & \textcolor{blue}{$2$} &  & \textcolor{blue}{$\underline{3}$} & & \textcolor{blue}{$2$} \\ \cline{3-8}
& \multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Right}}} & \textcolor{red}{$\underline{7}$} &  & \textcolor{red}{$2$} & & \textcolor{red}{$0$} & \\ 
& & \multicolumn{1}{|c}{} & \textcolor{blue}{$0$} &  & \textcolor{blue}{$5$} & & \textcolor{blue}{$\underline{7}$} \\ \cline{3-8}
\end{tabular}
\end{figure}

\columnsend

### Rationalizability II 

\columnsbegin

\column{.6\textwidth}

- We can **rationalize** other possibilities
- For example \textcolor{red}{Row} can rationalize playing \textcolor{red}{Left}
    - If she thinks \textcolor{blue}{Column} will play \textcolor{blue}{Right}, \textcolor{red}{Left} is her best response
    - \textcolor{blue}{Column} might play \textcolor{blue}{Right} because he thinks \textcolor{red}{Row} will play \textcolor{red}{Right}, it's his best response
- Similarly, we can **rationalize** many outcomes umder certain beliefs players hold

\column{.4\textwidth}

\begin{figure}
\centering
\begin{tabular}{llcc|cc|cc|}
& & \multicolumn{6}{c} {\textbf{\textcolor{blue}{Column}}} \\
& & \multicolumn{2}{c}{\textcolor{blue}{Left}} & \multicolumn{2}{c}{\textcolor{blue}{Middle}} & \multicolumn{2}{c}{\textcolor{blue}{Right}} \\ \cline{3-8}
\parbox[t]{2mm}{\multirow{6}{*}{\rotatebox[origin=c]{90}{\textbf{\textcolor{red}{Row}}}}}  & 
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Left}}} & \textcolor{red}{$0$} & & \textcolor{red}{$2$} & & \textcolor{red}{$\underline{7}$} & \\ 
&  &\multicolumn{1}{|c}{} & \textcolor{blue}{$\underline{7}$} & & \textcolor{blue}{$5$} & & \textcolor{blue}{$0$}\\ \cline{3-8}
&
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Middle}}} & \textcolor{red}{$5$} &  & \textcolor{red}{$\underline{3}$} & & \textcolor{red}{$5$} & \\ 
& & \multicolumn{1}{|c}{} & \textcolor{blue}{$2$} &  & \textcolor{blue}{$\underline{3}$} & & \textcolor{blue}{$2$} \\ \cline{3-8}
& \multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Right}}} & \textcolor{red}{$\underline{7}$} &  & \textcolor{red}{$2$} & & \textcolor{red}{$0$} & \\ 
& & \multicolumn{1}{|c}{} & \textcolor{blue}{$0$} &  & \textcolor{blue}{$5$} & & \textcolor{blue}{$\underline{7}$} \\ \cline{3-8}
\end{tabular}
\end{figure}

\columnsend

### Rationalizability III

\columnsbegin

\column{.6\textwidth}

- In this game, *any* of the 9 outcomes are **rationalizable** under certain beliefs!
- Each strategy for each player is a best response to *some* opposing strategy
    - Thus, each player *might* play any strategy if they believe their opponent will play the strategy it is a best response to

\column{.4\textwidth}

\begin{figure}
\centering
\begin{tabular}{llcc|cc|cc|}
& & \multicolumn{6}{c} {\textbf{\textcolor{blue}{Column}}} \\
& & \multicolumn{2}{c}{\textcolor{blue}{Left}} & \multicolumn{2}{c}{\textcolor{blue}{Middle}} & \multicolumn{2}{c}{\textcolor{blue}{Right}} \\ \cline{3-8}
\parbox[t]{2mm}{\multirow{6}{*}{\rotatebox[origin=c]{90}{\textbf{\textcolor{red}{Row}}}}}  & 
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Left}}} & \textcolor{red}{$0$} & & \textcolor{red}{$2$} & & \textcolor{red}{$\underline{7}$} & \\ 
&  &\multicolumn{1}{|c}{} & \textcolor{blue}{$\underline{7}$} & & \textcolor{blue}{$5$} & & \textcolor{blue}{$0$}\\ \cline{3-8}
&
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Middle}}} & \textcolor{red}{$5$} &  & \textcolor{red}{$\underline{3}$} & & \textcolor{red}{$5$} & \\ 
& & \multicolumn{1}{|c}{} & \textcolor{blue}{$2$} &  & \textcolor{blue}{$\underline{3}$} & & \textcolor{blue}{$2$} \\ \cline{3-8}
& \multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Right}}} & \textcolor{red}{$\underline{7}$} &  & \textcolor{red}{$2$} & & \textcolor{red}{$0$} & \\ 
& & \multicolumn{1}{|c}{} & \textcolor{blue}{$0$} &  & \textcolor{blue}{$5$} & & \textcolor{blue}{$\underline{7}$} \\ \cline{3-8}
\end{tabular}
\end{figure}

\columnsend

### Rationalizability IV

\columnsbegin

\column{.6\textwidth}

- Rationalizability can be used to find a solution
- Consider the game with different payoffs
- Identify strategies that are *never a best response* given any beliefs about opponent and eliminate them
- Neither player here has any strictly dominated strategies

\column{.4\textwidth}

\begin{figure}
\centering
\begin{tabular}{llcc|cc|cc|}
& & \multicolumn{6}{c} {\textbf{\textcolor{blue}{Column}}} \\
& & \multicolumn{2}{c}{\textcolor{blue}{Left}} & \multicolumn{2}{c}{\textcolor{blue}{Middle}} & \multicolumn{2}{c}{\textcolor{blue}{Right}} \\ \cline{3-8}
\parbox[t]{2mm}{\multirow{6}{*}{\rotatebox[origin=c]{90}{\textbf{\textcolor{red}{Row}}}}}  & 
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Left}}} & \textcolor{red}{$3$} & & \textcolor{red}{$0$} & & \textcolor{red}{$2$} & \\ 
&  &\multicolumn{1}{|c}{} & \textcolor{blue}{$2$} & & \textcolor{blue}{$3$} & & \textcolor{blue}{$0$}\\ \cline{3-8}
&
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Middle}}} & \textcolor{red}{$1$} &  & \textcolor{red}{$2$} & & \textcolor{red}{$1$} & \\ 
& & \multicolumn{1}{|c}{} & \textcolor{blue}{$3$} &  & \textcolor{blue}{$0$} & & \textcolor{blue}{$2$} \\ \cline{3-8}
& \multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Right}}} & \textcolor{red}{$2$} &  & \textcolor{red}{$4$} & & \textcolor{red}{$0$} & \\ 
& & \multicolumn{1}{|c}{} & \textcolor{blue}{$1$} &  & \textcolor{blue}{$3$} & & \textcolor{blue}{$2$} \\ \cline{3-8}
\end{tabular}
\end{figure}

\columnsend

### Rationalizability V

\columnsbegin

\column{.6\textwidth}

- Look for best responses
- Nash Equilibrium: (\textcolor{red}{Right}, \textcolor{blue}{Middle})
- For \textcolor{red}{Row}, \textcolor{red}{Middle} is never a best response
- For \textcolor{blue}{Column}, \textcolor{red}{Right} is never a best response
- Eliminating both 

\column{.4\textwidth}

\begin{figure}
\centering
\begin{tabular}{llcc|cc|cc|}
& & \multicolumn{6}{c} {\textbf{\textcolor{blue}{Column}}} \\
& & \multicolumn{2}{c}{\textcolor{blue}{Left}} & \multicolumn{2}{c}{\textcolor{blue}{Middle}} & \multicolumn{2}{c}{\textcolor{blue}{Right}} \\ \cline{3-8}
\parbox[t]{2mm}{\multirow{6}{*}{\rotatebox[origin=c]{90}{\textbf{\textcolor{red}{Row}}}}}  & 
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Left}}} & \textcolor{red}{$\underline{3}$} & & \textcolor{red}{$0$} & & \textcolor{red}{$\underline{2}$} & \\ 
&  &\multicolumn{1}{|c}{} & \textcolor{blue}{$2$} & & \textcolor{blue}{$\underline{3}$} & & \textcolor{blue}{$0$}\\ \cline{3-8}
&
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Middle}}} & \textcolor{red}{$1$} &  & \textcolor{red}{$2$} & & \textcolor{red}{$1$} & \\ 
& & \multicolumn{1}{|c}{} & \textcolor{blue}{$\underline{3}$} &  & \textcolor{blue}{$0$} & & \textcolor{blue}{$2$} \\ \cline{3-8}
& \multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Right}}} & \textcolor{red}{$2$} &  & \textcolor{red}{$\underline{4}$} & & \textcolor{red}{$0$} & \\ 
& & \multicolumn{1}{|c}{} & \textcolor{blue}{$1$} &  & \textcolor{blue}{$\underline{3}$} & & \textcolor{blue}{$2$} \\ \cline{3-8}
\end{tabular}
\end{figure}

\columnsend

### Rationalizability VI 

\columnsbegin

\column{.6\textwidth}

- We now see that for \textcolor{blue}{Column}, \textcolor{blue}{Left} is never a best response
- Delete it 

\column{.4\textwidth}

\begin{figure}
\centering
\begin{tabular}{llcc|cc|}
& & \multicolumn{4}{c} {\textbf{\textcolor{blue}{Column}}} \\
& & \multicolumn{2}{c}{\textcolor{blue}{Left}} & \multicolumn{2}{c}{\textcolor{blue}{Middle}} \\ \cline{3-6}
\parbox[t]{2mm}{\multirow{4}{*}{\rotatebox[origin=c]{90}{\textbf{\textcolor{red}{Row}}}}}  & 
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Left}}} & \textcolor{red}{$\underline{3}$} & & \textcolor{red}{$0$} &  \\ 
&  &\multicolumn{1}{|c}{} & \textcolor{blue}{$2$} & & \textcolor{blue}{$\underline{3}$} \\ \cline{3-6}
& \multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Right}}} & \textcolor{red}{$2$} &  & \textcolor{red}{$\underline{4}$} &  \\ 
& & \multicolumn{1}{|c}{} & \textcolor{blue}{$1$} &  & \textcolor{blue}{$\underline{3}$}  \\ \cline{3-6}
\end{tabular}
\end{figure}

\columnsend

### Rationalizability VII 

\columnsbegin

\column{.6\textwidth}

- Finally, for \textcolor{red}{Row}, \textcolor{red}{Left} is never a best response
- We are left with the Nash equilibrium 

\column{.4\textwidth}

\begin{figure}
\centering
\begin{tabular}{llcc|}
& & \multicolumn{2}{c} {\textbf{\textcolor{blue}{Column}}} \\
& & \multicolumn{2}{c}{\textcolor{blue}{Middle}} \\ \cline{3-4}
\parbox[t]{2mm}{\multirow{4}{*}{\rotatebox[origin=c]{90}{\textbf{\textcolor{red}{Row}}}}}  & 
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Left}}}  & \textcolor{red}{$0$} &  \\ 
&  &\multicolumn{1}{|c}{} & \textcolor{blue}{$\underline{3}$} \\ \cline{3-4}
& \multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Right}}} &  \textcolor{red}{$\underline{4}$} &  \\ 
& & \multicolumn{1}{|c}{}  & \textcolor{blue}{$\underline{3}$}  \\ \cline{3-4}
\end{tabular}
\end{figure}

\columnsend

### Rationalizability VIII

\columnsbegin

\column{.6\textwidth}

- Again, consider the game from before
- For \textcolor{red}{Row}:
    - \textcolor{red}{Left} is rationalizable if \textcolor{red}{Row} believes \textcolor{blue}{Column} plays \textcolor{blue}{Right}
    - \textcolor{red}{Middle} is rationalizable if \textcolor{red}{Row} believes \textcolor{blue}{Column} plays \textcolor{blue}{Middle}
    - \textcolor{red}{Right} is rationalizable if \textcolor{red}{Row} believes \textcolor{blue}{Right} plays \textcolor{blue}{Left}

\column{.4\textwidth}

\begin{figure}
\centering
\begin{tabular}{llcc|cc|cc|}
& & \multicolumn{6}{c} {\textbf{\textcolor{blue}{Column}}} \\
& & \multicolumn{2}{c}{\textcolor{blue}{Left}} & \multicolumn{2}{c}{\textcolor{blue}{Middle}} & \multicolumn{2}{c}{\textcolor{blue}{Right}} \\ \cline{3-8}
\parbox[t]{2mm}{\multirow{6}{*}{\rotatebox[origin=c]{90}{\textbf{\textcolor{red}{Row}}}}}  & 
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Left}}} & \textcolor{red}{$0$} & & \textcolor{red}{$2$} & & \textcolor{red}{$\underline{7}$} & \\ 
&  &\multicolumn{1}{|c}{} & \textcolor{blue}{$\underline{7}$} & & \textcolor{blue}{$5$} & & \textcolor{blue}{$0$}\\ \cline{3-8}
&
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Middle}}} & \textcolor{red}{$5$} &  & \textcolor{red}{$\underline{3}$} & & \textcolor{red}{$5$} & \\ 
& & \multicolumn{1}{|c}{} & \textcolor{blue}{$2$} &  & \textcolor{blue}{$\underline{3}$} & & \textcolor{blue}{$2$} \\ \cline{3-8}
& \multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Right}}} & \textcolor{red}{$\underline{7}$} &  & \textcolor{red}{$2$} & & \textcolor{red}{$0$} & \\ 
& & \multicolumn{1}{|c}{} & \textcolor{blue}{$0$} &  & \textcolor{blue}{$5$} & & \textcolor{blue}{$\underline{7}$} \\ \cline{3-8}
\end{tabular}
\end{figure}

\columnsend

### Rationalizability VIII

\columnsbegin

\column{.6\textwidth}

- Again, consider the game from before
- For \textcolor{blue}{Column}:
    - \textcolor{blue}{Left} is rationalizable if \textcolor{blue}{Column} believes \textcolor{red}{Row} plays \textcolor{red}{Left}
    - \textcolor{blue}{Middle} is rationalizable if \textcolor{blue}{Column} believes \textcolor{red}{Row} plays \textcolor{red}{Middle}
    - \textcolor{blue}{Right} is rationalizable if \textcolor{blue}{Column} believes \textcolor{red}{Row} plays \textcolor{red}{Right}
- All 9 outcomes are rationalizable!
    - But again, we saw a unique Nash equilibrium (using best-response): (\textcolor{red}{Middle}, \textcolor{blue}{Middle})

\column{.4\textwidth}

\begin{figure}
\centering
\begin{tabular}{llcc|cc|cc|}
& & \multicolumn{6}{c} {\textbf{\textcolor{blue}{Column}}} \\
& & \multicolumn{2}{c}{\textcolor{blue}{Left}} & \multicolumn{2}{c}{\textcolor{blue}{Middle}} & \multicolumn{2}{c}{\textcolor{blue}{Right}} \\ \cline{3-8}
\parbox[t]{2mm}{\multirow{6}{*}{\rotatebox[origin=c]{90}{\textbf{\textcolor{red}{Row}}}}}  & 
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Left}}} & \textcolor{red}{$0$} & & \textcolor{red}{$2$} & & \textcolor{red}{$\underline{7}$} & \\ 
&  &\multicolumn{1}{|c}{} & \textcolor{blue}{$\underline{7}$} & & \textcolor{blue}{$5$} & & \textcolor{blue}{$0$}\\ \cline{3-8}
&
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Middle}}} & \textcolor{red}{$5$} &  & \textcolor{red}{$\underline{3}$} & & \textcolor{red}{$5$} & \\ 
& & \multicolumn{1}{|c}{} & \textcolor{blue}{$2$} &  & \textcolor{blue}{$\underline{3}$} & & \textcolor{blue}{$2$} \\ \cline{3-8}
& \multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Right}}} & \textcolor{red}{$\underline{7}$} &  & \textcolor{red}{$2$} & & \textcolor{red}{$0$} & \\ 
& & \multicolumn{1}{|c}{} & \textcolor{blue}{$0$} &  & \textcolor{blue}{$5$} & & \textcolor{blue}{$\underline{7}$} \\ \cline{3-8}
\end{tabular}
\end{figure}

\columnsend

### Rationalizability IX

\columnsbegin

\column{.6\textwidth}

- All 9 outcomes are rationalizable (under certain beliefs)!
- But again, a unique Nash equilibrium (using best-response): (\textcolor{red}{Middle}, \textcolor{blue}{Middle})
- Rationalizability process (eliminating strategies that are never a best response) may not lead us to Nash equilibrium
- All Nash equilibria are rationalizable, but not all rationalizable equilibria are Nash equilibria

\column{.4\textwidth}

\begin{figure}
\centering
\begin{tabular}{llcc|cc|cc|}
& & \multicolumn{6}{c} {\textbf{\textcolor{blue}{Column}}} \\
& & \multicolumn{2}{c}{\textcolor{blue}{Left}} & \multicolumn{2}{c}{\textcolor{blue}{Middle}} & \multicolumn{2}{c}{\textcolor{blue}{Right}} \\ \cline{3-8}
\parbox[t]{2mm}{\multirow{6}{*}{\rotatebox[origin=c]{90}{\textbf{\textcolor{red}{Row}}}}}  & 
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Left}}} & \textcolor{red}{$0$} & & \textcolor{red}{$2$} & & \textcolor{red}{$\underline{7}$} & \\ 
&  &\multicolumn{1}{|c}{} & \textcolor{blue}{$\underline{7}$} & & \textcolor{blue}{$5$} & & \textcolor{blue}{$0$}\\ \cline{3-8}
&
\multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Middle}}} & \textcolor{red}{$5$} &  & \textcolor{red}{$\underline{3}$} & & \textcolor{red}{$5$} & \\ 
& & \multicolumn{1}{|c}{} & \textcolor{blue}{$2$} &  & \textcolor{blue}{$\underline{3}$} & & \textcolor{blue}{$2$} \\ \cline{3-8}
& \multicolumn{1}{c|}{\multirow{2}{*}{\textcolor{red}{Right}}} & \textcolor{red}{$\underline{7}$} &  & \textcolor{red}{$2$} & & \textcolor{red}{$0$} & \\ 
& & \multicolumn{1}{|c}{} & \textcolor{blue}{$0$} &  & \textcolor{blue}{$5$} & & \textcolor{blue}{$\underline{7}$} \\ \cline{3-8}
\end{tabular}
\end{figure}

\columnsend
